---
title: Modeling with Data in the Tidyverse
author: ''
date: '2021-08-16'
slug: []
categories: []
tags:
  - data_wrangling
  - Classification
  - Regression
---

```{r}
# Load packages
library(moderndive)
library(ggplot2)
library(tidyverse)
```

## 1-1 Log10 transformation of house size

You just saw that the predictor variable sqft_living is right-skewed and hence a log base 10 transformation is warranted to unskew it. Just as we transformed the outcome variable price to create log10_price in the video, let's do the same for sqft_living.

```{r}

# Plot the histogram
ggplot(house_prices, aes(x = sqft_living)) +
  geom_histogram(bins = 30) +
  labs(x="Size (sq.feet)", y="count")
```


```{r}
# Add log10_size
house_prices_2 <- house_prices %>%
  mutate(log10_size = log10(sqft_living))

# Plot the histogram  
ggplot(house_prices_2, aes(x = log10_size)) +
  geom_histogram() +
  labs(x = "log10 size", y = "count")
```
## 1-2 adding jitter

The researchers in the UT Austin created a "beauty score" by asking a panel of 6 students to rate the "beauty" of all 463 instructors. They were interested in studying any possible impact of "beauty" of teaching evaluation scores. Let's do an EDA of this variable and its relationship with teaching score.

From now on, assume that ggplot2, dplyr, and moderndive are all available in your workspace unless you're told otherwise.

```{r}
# Scatterplot
ggplot(evals, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "beauty score", y = "teaching score")
```

```{r}
# Jitter plot
ggplot(evals, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  labs(x = "beauty score", y = "teaching score")
```

