---
title: Unsupervised Learning in R
author: ''
date: '2021-08-26'
slug: []
categories: []
tags:
  - Unsupervised Learning
---

```{r}
x <- read.csv("x.csv")
library(tidyverse)
x %>% ggplot(aes(V1,V2))+geom_point(alpha=0.9, shape="x")
```

## 1-4 k-means clustering

We have created some two-dimensional data and stored it in a variable called x in your workspace. The scatter plot on the right is a visual representation of the data.

In this exercise, your task is to create a k-means model of the x data using 3 clusters, then to look at the structure of the resulting model using the summary() function.

```{r}
# Create the k-means model: km.out
km.out <- kmeans(x, centers = 3, nstart = 20)

# Inspect the result
summary(km.out)
```

## 1-5 Results of kmeans()

The kmeans() function produces several outputs. In the video, we discussed one output of modeling, the cluster membership.

In this exercise, you will access the cluster component directly. This is useful anytime you need the cluster membership for each observation of the data used to build the clustering model. A future exercise will show an example of how this cluster membership might be used to help communicate the results of k-means modeling.

k-means models also have a print method to give a human friendly output of basic modeling results. This is available by using print() or simply typing the name of the model.

```{r}
# Print the cluster membership component of the model

print(km.out$cluster)
# Print the km.out object
print(km.out)

```

## 1-6 Visualizing and interpreting results of kmeans()

One of the more intuitive ways to interpret the results of k-means models is by plotting the data as a scatter plot and using color to label the samples' cluster membership. In this exercise, you will use the standard plot() function to accomplish this.

To create a scatter plot, you can pass data with two features (i.e. columns) to plot() with an extra argument col = km.out$cluster, which sets the color of each point in the scatter plot according to its cluster membership.

```{r}
# Scatter plot of x
x %>%
  ggplot(aes(V1,V2))+geom_point(color=km.out$cluster)
```

## 1-8 

```{r}

```

## 1-9 

```{r}

```

## 1-11 

```{r}

```

