---
title: Unsupervised Learning in R
author: ''
date: '2021-08-26'
slug: []
categories: []
tags:
  - Unsupervised Learning
---

```{r}
library(tidyverse)
x <- read.csv("x.csv") %>%select(c(V1,V2))
pokemon<-read.csv("pokemon.csv")%>%select(-c(X))
x2 <- read.csv("x2.csv")

x %>% ggplot(aes(V1,V2))+geom_point(alpha=0.9, shape="x", size=2)
```

## 1-4 k-means clustering

We have created some two-dimensional data and stored it in a variable called x in your workspace. Above scatter plot is a visual representation of the data.

In this exercise, your task is to create a k-means model of the x data using 3 clusters, then to look at the structure of the resulting model using the summary() function.

```{r}
# Create the k-means model: km.out
km.out <- kmeans(x, centers = 3, nstart = 200)

# Inspect the result
summary(km.out)
```

## 1-5 Results of kmeans()

The kmeans() function produces several outputs. In the video, we discussed one output of modeling, the cluster membership.

In this exercise, you will access the cluster component directly. This is useful anytime you need the cluster membership for each observation of the data used to build the clustering model. A future exercise will show an example of how this cluster membership might be used to help communicate the results of k-means modeling.

k-means models also have a print method to give a human friendly output of basic modeling results. This is available by using print() or simply typing the name of the model.

```{r}
# Print the cluster membership component of the model

print(km.out$cluster)
# Print the km.out object
print(km.out)

```

## 1-6 Visualizing and interpreting results of kmeans()

One of the more intuitive ways to interpret the results of k-means models is by plotting the data as a scatter plot and using color to label the samples' cluster membership. In this exercise, you will use the standard plot() function to accomplish this.

To create a scatter plot, you can pass data with two features (i.e. columns) to plot() with an extra argument col = km.out$cluster, which sets the color of each point in the scatter plot according to its cluster membership.

```{r}
# Scatter plot of x
x %>%
  ggplot(aes(V1,V2))+geom_point(color=km.out$cluster)
```

## 1-8 Handling random algorithms

In the video, you saw how kmeans() randomly initializes the centers of clusters. This random initialization can result in assigning observations to different cluster labels. Also, the random initialization can result in finding different local minima for the k-means algorithm. This exercise will demonstrate both results.

At the top of each plot, the measure of model quality—total within cluster sum of squares error—will be plotted. Look for the model(s) with the lowest error to find models with the better model results.

Because kmeans() initializes observations to random clusters, it is important to set the random number generator seed for reproducibility.

```{r warning=FALSE}
# Set up 2 x 3 plotting grid
par(mfrow = c(2, 3))

# Set seed
set.seed(1)

for(i in 1:6) {
  # Run kmeans() on x with three clusters and one start
  km.out <- kmeans(x, centers = 3, nstart = 1)
  
  # Plot clusters
  plot(x, col = km.out$cluster, 
       main = km.out$tot.withinss)
}
```


```{r}
# Set up 2 x 3 plotting grid
par(mfrow = c(2, 3))

# Set seed
set.seed(1)

for(i in 1:6) {
  # Run kmeans() on x with three clusters and one start
  km.out <- kmeans(x, centers = 3, nstart = 1)
  
  # Plot clusters
  plot(x, col = km.out$cluster, 
       main = km.out$tot.withinss)
}
```


## 1-9 Selecting number of clusters

The k-means algorithm assumes the number of clusters as part of the input. If you know the number of clusters in advance (e.g. due to certain business constraints) this makes setting the number of clusters easy. However, as you saw in the video, if you do not know the number of clusters and need to determine it, you will need to run the algorithm multiple times, each time with a different number of clusters. From this, you can observe how a measure of model quality changes with the number of clusters.

In this exercise, you will run kmeans() multiple times to see how model quality changes as the number of clusters changes. Plots displaying this information help to determine the number of clusters and are often referred to as scree plots.

The ideal plot will have an elbow where the quality measure improves more slowly as the number of clusters increases. This indicates that the quality of the model is no longer improving substantially as the model complexity (i.e. number of clusters) increases. In other words, the elbow indicates the number of clusters inherent in the data.

```{r}
# Initialize total within sum of squares error: wss
wss <- 0

# For 1 to 15 cluster centers
for (i in 1:15) {
  km.out <- kmeans(x, centers = i, nstart = 20)
  # Save total within sum of squares to wss variable
  wss[i] <- km.out$tot.withinss
}

# Plot total within sum of squares vs. number of clusters
plot(1:15, wss, type = "b", 
     xlab = "Number of Clusters", 
     ylab = "Within groups sum of squares")

# Set k equal to the number of clusters corresponding to the elbow location
k <- 2  # 3 is probably OK, too
```

## 1-11 Practical matters: working with real data

Dealing with real data is often more challenging than dealing with synthetic data. Synthetic data helps with learning new concepts and techniques, but the next few exercises will deal with data that is closer to the type of real data you might find in your professional or academic pursuits.

The first challenge with the Pokemon data is that there is no pre-determined number of clusters. You will determine the appropriate number of clusters, keeping in mind that in real data the elbow in the scree plot might be less of a sharp elbow than in synthetic data. Use your judgement on making the determination of the number of clusters.

The second part of this exercise includes plotting the outcomes of the clustering on two dimensions, or features, of the data. These features were chosen somewhat arbitrarily for this exercise. Think about how you would use plotting and clustering to communicate interesting groups of Pokemon to other people.

An additional note: this exercise utilizes the iter.max argument to kmeans(). As you've seen, kmeans() is an iterative algorithm, repeating over and over until some stopping criterion is reached. The default number of iterations for kmeans() is 10, which is not enough for the algorithm to converge and reach its stopping criterion, so we'll set the number of iterations to 50 to overcome this issue. To see what happens when kmeans() does not converge, try running the example with a lower number of iterations (e.g. 3). This is another example of what might happen when you encounter real data and use real cases.


```{r}
# Initialize total within sum of squares error: wss
wss <- 0

# Look over 1 to 15 possible clusters
for (i in 1:15) {
  # Fit the model: km.out
  km.out <- kmeans(pokemon, centers = i, nstart = 20, iter.max = 50)
  # Save the within cluster sum of squares
  wss[i] <- km.out$tot.withinss
}

# Produce a scree plot
plot(1:15, wss, type = "b", 
     xlab = "Number of Clusters", 
     ylab = "Within groups sum of squares")

# Select number of clusters
k <- 3

# Build model with k clusters: km.out
km.out <- kmeans(pokemon, centers = k, nstart = 20, iter.max = 50)

# View the resulting model
km.out

# Plot of Defense vs. Speed by cluster membership
plot(pokemon[, c("Defense", "Speed")],
     col = km.out$cluster,
     main = paste("k-means clustering of Pokemon with", k, "clusters"),
     xlab = "Defense", ylab = "Speed")
```


## 2-2 Hierarchical clustering with results

In this exercise, you will create your first hierarchical clustering model using the hclust() function.

We have created some data that has two dimensions and placed it in a variable called x2. Your task is to create a hierarchical clustering model of x2. Remember from the video that the first step to hierarchical clustering is determining the similarity between observations, which you will do with the dist() function.

You will look at the structure of the resulting model using the summary() function.

```{r}
# Create hierarchical clustering model: hclust.out
hclust.out <- hclust(d=dist(x2) )

# Inspect the result
summary(hclust.out)
```

## 2-5 Cutting the tree

Remember from the video that cutree() is the R function that cuts a hierarchical model. The h and k arguments to cutree() allow you to cut the tree based on a certain height h or a certain number of clusters k.

In this exercise, you will use cutree() to cut the hierarchical model you created earlier based on each of these two criteria.

```{r}
# Cut by height
cutree(hclust.out, h = 7)

# Cut by number of clusters
cutree(hclust.out, k = 3)
```

## 2-7 Linkage methods

In this exercise, you will produce hierarchical clustering models using different linkages and plot the dendrogram for each, observing the overall structure of the trees.

You'll be asked to interpret the results in the next exercise.

```{r}
# Cluster using complete linkage: hclust.complete
hclust.complete <- hclust(dist(x2), method = "complete")

# Cluster using average linkage: hclust.average
hclust.average <- hclust(dist(x2), method = "average")

# Cluster using single linkage: hclust.single
hclust.single <- hclust(dist(x2), method = "single")

# Plot dendrogram of hclust.complete
plot(hclust.complete, main = "Complete")

# Plot dendrogram of hclust.average
plot(hclust.average, main = "Average")

# Plot dendrogram of hclust.single
plot(hclust.single, main = "Single")
```

## 2-9 Practical matters: scaling

Recall from the video that clustering real data may require scaling the features if they have different distributions. So far in this chapter, you have been working with synthetic data that did not need scaling.

In this exercise, you will go back to working with "real" data, the pokemon dataset introduced in the first chapter. You will observe the distribution (mean and standard deviation) of each feature, scale the data accordingly, then produce a hierarchical clustering model using the complete linkage method.

```{r}
# View column means
colMeans(pokemon)

# View column standard deviations
 apply(pokemon, 2, sd)

# Scale the data
pokemon.scaled <- scale(pokemon)

# Create hierarchical clustering model: hclust.pokemon
hclust.pokemon <- hclust(dist(pokemon.scaled), method = "complete")
```

## 2-10 Comparing kmeans() and hclust()

Comparing k-means and hierarchical clustering, you'll see the two methods produce different cluster memberships. This is because the two algorithms make different assumptions about how the data is generated. In a more advanced course, we could choose to use one model over another based on the quality of the models' assumptions, but for now, it's enough to observe that they are different.

This exercise will have you compare results from the two models on the pokemon dataset to see how they differ.

```{r}
# Apply cutree() to hclust.pokemon: cut.pokemon
cut.pokemon <- cutree(hclust.pokemon, k = 3)

# Compare methods
table(km.out$cluster, cut.pokemon)
```

## 3-2 PCA using prcomp()

In this exercise, you will create your first PCA model and observe the diagnostic results.

We have loaded the Pokemon data from earlier, which has four dimensions, and placed it in a variable called pokemon. Your task is to create a PCA model of the data, then to inspect the resulting model using the summary() function.

hint: (for unknown reason) do not consider SpecialAttack and SpecialDefence variables.

```{r}
# Perform scaled PCA: pr.out
pr.out <- prcomp(x = pokemon[-c(4,5)], scale = TRUE, center = TRUE) 

pr.out
# Inspect model output
summary(pr.out)
```
```{r}
apply(pokemon[-c(4,5)],2,sd)
print("----------------------------------")
apply(pokemon[-c(4,5)],2,mean)
print("----------------------------------")
apply(scale(pokemon[-c(4,5)]),2,sd)
apply(scale(pokemon[-c(4,5)]),2,mean)
```


Q: What is the minimum number of principal components that are required to describe at least 75% of the cumulative variance in this dataset?

```{r}
summary(pr.out)
```

A: 2

## 3-7 Interpreting biplots (1)

As stated in the video, the biplot() function plots both the principal components loadings and the mapping of the observations to their first two principal component values. The next couple of exercises will check your interpretation of the biplot() visualization.

Using the biplot() of the pr.out model, which two original variables have approximately the same loadings in the first two principal components?

```{r}
biplot(pr.out)
```

A: Attack and HitPoints

Q: which two Pokemon are the least similar in terms of the second principal component?

A: Kadabra and Torkoal


## 3-8 Variance explained

The second common plot type for understanding PCA models is a scree plot. A scree plot shows the variance explained as the number of principal components increases. Sometimes the cumulative variance explained is plotted as well.

In this and the next exercise, you will prepare data from the pr.out model you created at the beginning of the chapter for use in a scree plot. Preparing the data for plotting is required because there is not a built-in function in R to create this type of plot.

```{r}
# Variability of each principal component: pr.var
pr.var <- pr.out$sdev^2

# Variance explained by each principal component: pve
pve <-  pr.var / sum(pr.var)
```

## 3-9 Visualize variance explained

Now you will create a scree plot showing the proportion of variance explained by each principal component, as well as the cumulative proportion of variance explained.

Recall from the video that these plots can help to determine the number of principal components to retain. One way to determine the number of principal components to retain is by looking for an elbow in the scree plot showing that as the number of principal components increases, the rate at which variance is explained decreases substantially. In the absence of a clear elbow, you can use the scree plot as a guide for setting a threshold.

```{r}
# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component",
     ylab = "Proportion of Variance Explained",
     ylim = c(0, 1), type = "b")

# Plot cumulative proportion of variance explained
plot(cumsum(pve), xlab = "Principal Component",
     ylab = "Cumulative Proportion of Variance Explained",
     ylim = c(0, 1), type = "b")
```

## 3-11 Practical issues: scaling

You saw in the video that scaling your data before doing PCA changes the results of the PCA modeling. Here, you will perform PCA with and without scaling, then visualize the results using biplots.

Sometimes scaling is appropriate when the variances of the variables are substantially different. This is commonly the case when variables have different units of measurement, for example, degrees Fahrenheit (temperature) and miles (distance). Making the decision to use scaling is an important step in performing a principal component analysis.

```{r}
# Mean of each variable
colMeans(pokemon)

# Standard deviation of each variable
apply(pokemon, 2, sd)

# PCA model with scaling: pr.with.scaling
pr.with.scaling <- prcomp(pokemon, scale = TRUE)

# PCA model without scaling: pr.without.scaling
pr.without.scaling <- prcomp(pokemon, scale = FALSE)

# Create biplots of both for comparison
biplot(pr.with.scaling)
biplot(pr.without.scaling)
```

## 4-2 Preparing the data
Unlike prior chapters, where we prepared the data for you for unsupervised learning, the goal of this chapter is to step you through a more realistic and complete workflow.

Recall from the video that the first step is to download and prepare the data.

Use read.csv() function to download the CSV (comma-separated values) file containing the data from the URL provided. Assign the result to wisc.df.


tip for last part: set a vector called diagnosis to be 1 if a diagnosis is malignant ("M") and 0 otherwise. Note that R coerces TRUE to 1 and FALSE to 0.

```{r}
url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1903/datasets/WisconsinCancer.csv"

# Download the data: wisc.df
wisc.df <- read.csv(url)

# Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[3:32])

# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id

# Create diagnosis vector
diagnosis <- as.numeric(wisc.df$diagnosis == "M")
```



## 4-4 Performing PCA

The next step in your analysis is to perform PCA on wisc.data.

You saw in the last chapter that it's important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data:

The input variables use different units of measurement.
The input variables have significantly different variances.


```{r}
# Check column means and standard deviations
colMeans(wisc.data)
apply(wisc.data,2 , sd)

# Execute PCA, scaling if appropriate: wisc.pr
wisc.pr <- prcomp(x = wisc.data, scale = TRUE, center = TRUE)

# Look at summary of results
summary(wisc.pr)
```

## 4-5 Interpreting PCA results

Now you'll use some visualizations to better understand your PCA model. You were introduced to one of these visualizations, the biplot, in an earlier chapter.

You'll run into some common challenges with using biplots on real-world data containing a non-trivial number of observations and variables, then you'll look at some alternative visualizations. You are encouraged to experiment with additional visualizations before moving on to the next exercise.

```{r}
# Create a biplot of wisc.pr
biplot(wisc.pr)

# Scatter plot observations by components 1 and 2
plot(wisc.pr$x[, c(1, 2)], col = (diagnosis + 1), 
     xlab = "PC1", ylab = "PC2")

# Repeat for components 1 and 3
plot(wisc.pr$x[, c(1, 3)], col = (diagnosis + 1), 
     xlab = "PC1", ylab = "PC3")

# Do additional data exploration of your choosing below (optional)

```

## 4-6 Variance explained

In this exercise, you will produce scree plots showing the proportion of variance explained as the number of principal components increases. The data from PCA must be prepared for these plots, as there is not a built-in function in R to create them directly from the PCA model.

As you look at these plots, ask yourself if there's an elbow in the amount of variance explained that might lead you to pick a natural number of principal components. If an obvious elbow does not exist, as is typical in real-world datasets, consider how else you might determine the number of principal components to retain based on the scree plot.

```{r}
# Set up 1 x 2 plotting grid
par(mfrow = c(1, 2))

# Calculate variability of each component
pr.var <- wisc.pr$sdev^2

# Variance explained by each principal component: pve
pve <- pr.var / sum(pr.var)

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "b")

# Plot cumulative proportion of variance explained
plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "b")
```

## 4-9 Hierarchical clustering of case data

The goal of this exercise is to do hierarchical clustering of the observations. Recall from Chapter 2 that this type of clustering does not assume in advance the number of natural groups that exist in the data.

As part of the preparation for hierarchical clustering, distance between all pairs of observations are computed. Furthermore, there are different ways to link clusters together, with single, complete, and average being the most common linkage methods.


```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data)

# Calculate the (Euclidean) distances: data.dist
data.dist <- dist(data.scaled)

# Create a hierarchical clustering model: wisc.hclust
wisc.hclust <- hclust(data.dist, method = "complete")
```

## 4-10 Results of hierarchical clustering
Let's use the hierarchical clustering model you just created to determine a height (or distance between clusters) where a certain number of clusters exists. The variables you created before—wisc.data, diagnosis, wisc.pr, pve, and wisc.hclust—are all available in your workspace.

Using the plot() function, what is the height at which the clustering model has 4 clusters?

```{r}
plot(wisc.hclust)
```
A: 20

## 4-11 Selecting number of clusters

In this exercise, you will compare the outputs from your hierarchical clustering model to the actual diagnoses. Normally when performing unsupervised learning like this, a target variable isn't available. We do have it with this dataset, however, so it can be used to check the performance of the clustering model.

When performing supervised learning—that is, when you're trying to predict some target variable of interest and that target variable is available in the original data—using clustering to create new features may or may not improve the performance of the final model. This exercise will help you determine if, in this case, hierarchical clustering provides a promising new feature.


```{r}
# Cut tree so that it has 4 clusters: wisc.hclust.clusters
wisc.hclust.clusters <- cutree(wisc.hclust, k = 4)

# Compare cluster membership to actual diagnoses
table(wisc.hclust.clusters, diagnosis)
```

## 4-12 k-means clustering and comparing results

As you now know, there are two main types of clustering: hierarchical and k-means.

In this exercise, you will create a k-means clustering model on the Wisconsin breast cancer data and compare the results to the actual diagnoses and the results of your hierarchical clustering model. Take some time to see how each clustering model performs in terms of separating the two diagnoses and how the clustering models compare to each other.



```{r}
# Create a k-means model on wisc.data: wisc.km
wisc.km <- kmeans(scale(wisc.data), centers = 2, nstart = 20)

# Compare k-means to actual diagnoses
table(wisc.km$cluster , diagnosis)

# Compare k-means to hierarchical clustering
table(wisc.km$cluster, wisc.hclust.clusters)
```

## 4-13 Clustering on PCA results

In this final exercise, you will put together several steps you used earlier and, in doing so, you will experience some of the creativity that is typical in unsupervised learning.

Recall from earlier exercises that the PCA model required significantly fewer features to describe 80% and 95% of the variability of the data. In addition to normalizing data and potentially avoiding overfitting, PCA also uncorrelates the variables, sometimes improving the performance of other modeling techniques.

Let's see if PCA improves or degrades the performance of hierarchical clustering.


```{r}
# Create a hierarchical clustering model: wisc.pr.hclust
wisc.pr.hclust <- hclust(dist(wisc.pr$x[, 1:7]), method = "complete")

# Cut model into 4 clusters: wisc.pr.hclust.clusters
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k = 4)

# Compare to actual diagnoses
table(diagnosis, wisc.pr.hclust.clusters)

# Compare to k-means and hierarchical
table(diagnosis, wisc.hclust.clusters)
table(diagnosis, wisc.km$cluster)
```

