---
title: Modeling with tidymodels in R
author: ''
date: '2021-07-20'
slug: []
categories:
  - tidymodels
tags:
  - Machine Learning
---

```{r}
library(tidymodels)
library(tidyverse)
```

tidymodels is a collection of machine learning packages designed to simplify the machine learning workflow in R.

```{r}
home_sales <- read_csv("home_sales.csv")
```

## 1-3 Creating training and test datasets

The rsample package is designed to create training and test datasets. Creating a test dataset is important for estimating how a trained model will likely perform on new data. It also guards against overfitting, where a model memorizes patterns that exist only in the training data and performs poorly on new data.

This data contains information on homes sold in the Seattle, Washington area between 2015 and 2016.
The outcome variable in this data is selling_price.

```{r}
# Create a data split object

# Create a data split object
home_split <- initial_split(home_sales, 
                            prop = 0.7, 
                            strata = selling_price)

# Create the training data
home_training <- home_split %>%
  training()

# Create the test data
home_test <- home_split %>% 
  testing()

# Check number of rows in each dataset
nrow(home_training)
nrow(home_test)

```

## 1-4 Distribution of outcome variable values

Stratifying by the outcome variable when generating training and test datasets ensures that the outcome variable values have a similar range in both datasets.

Since the original data is split at random, stratification avoids placing all the expensive homes in home_sales into the test dataset, for example. In this case, your model would most likely perform poorly because it was trained on less expensive homes.

In this exercise, you will calculate summary statistics for the selling_price variable in the training and test datasets. The home_training and home_test tibbles have been loaded from the previous exercise.

```{r}
# Distribution of selling_price in training data
home_training %>% 
  summarize(min_sell_price = min(selling_price),
            max_sell_price = max(selling_price),
            mean_sell_price = mean(selling_price),
            sd_sell_price = sd(selling_price))





# Distribution of selling_price in test data
home_test %>% 
  summarize(min_sell_price = min(selling_price),
            max_sell_price = max(selling_price),
            mean_sell_price = mean(selling_price),
            sd_sell_price = sd(selling_price))
```

## 1-6 Fitting a linear regression model

The parsnip package provides a unified syntax for the model fitting process in R.

With parsnip, it is easy to define models using the various packages, or engines, that exist in the R ecosystem.

In this exercise, you will define a parsnip linear regression object and train your model to predict selling_price using home_age and sqft_living as predictor variables from the home_sales data.

The home_training and home_test tibbles that you created in the previous lesson have been loaded into this session.

```{r}

# Specify a linear regression model, linear_model
linear_model <- linear_reg() %>% 
  # Set the model engine
  set_engine('lm') %>% 
  # Set the model mode
  set_mode('regression')

# Train the model with the training data
lm_fit <- linear_model %>% 
  fit(selling_price~home_age+sqft_living,
      data = home_training)

# Print lm_fit to view model information
lm_fit


tidy(lm_fit)
```

## 1-8 Predicting home selling prices

After fitting a model using the training data, the next step is to use it to make predictions on the test dataset. The test dataset acts as a new source of data for the model and will allow you to evaluate how well it performs.

Before you can evaluate model performance, you must add your predictions to the test dataset.

In this exercise, you will use your trained model, lm_fit, to predict selling_price in the home_test dataset.

Your trained model, lm_fit, as well as the test dataset, home_test have been loaded into your session.