---
title: 'Assignment A02: Classification Models'
author: Ramin Ala
date: '2021-07-15'
slug: []
categories: []
tags:
  - Assignment
  - Classification
---
***Ramin Ala***


```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)

bank <- read_csv("bank.csv") %>%
  select(c(age, job, marital, education, default, balance, housing, loan, y)) %>%
  mutate(result=ifelse(y == 'yes', 1, 0))
 

```

## Executive Summary

The data is related with direct marketing campaigns (based on phone calls) of a Portuguese banking institution. This really is a **supervised question** because it has an outcome (y), which defined as if the product (bank term deposit) would be (or not) subscribed, "yes" means subscribed, and "no" means did not subscribe. This dataset consists of **`r nrow(bank)` observations**.

In addition to the outcome classification variable, there are 16 variables. Eight variables are related to statistical information about current and previous campaigns, and they seems irrelevant to  dependent variable. So, there are eight determinent variables Including:

Customer's ***age*** and ***balance*** (numerical variables),

Customer's ***job***, ***marital*** status, ***education***  (categorical variables)
 
***housing loan***, ***personal loan***, and ***default status*** (binary variables)  

Information regarding the reference of dataset is available at:

http://www3.dsi.uminho.pt/pcortez/bib/2011-esm-1.txt

## Univariate Analysis of numerical variables

## *age*

```{r, echo=FALSE, message=FALSE, warning=FALSE}
bank %>%
  ggplot(aes(age)) +
  geom_density(fill = "green", alpha = 0.4, size=0.01) +
   scale_x_continuous(breaks = seq(15, 90, by = 5))

summary(bank$age) %>%
  tidy() 
```

Above plot and table give an impression on the age of examined group. The minimum, maximum and average ages are 19, 87, and 41 years respectively. Majority of observations are concentrated at the age of between 30 to 35 years. This information would not provide any insight on the possible outcome.


## *balance*

```{r, echo=FALSE, message=FALSE, warning=FALSE}
bank %>%
  filter(between(balance,-4000,10000)) %>%
  ggplot(aes(balance)) +
  geom_histogram(fill = "red", alpha = 0.4, size=1) +
   scale_x_continuous(breaks = seq(-4000, 10000, by = 1000))

summary(bank$balance) %>%
  tidy()
```

Balance is generally between 0 to $500, however, amounts as high as $71000 is available in data. In the above plot, Only range from $-4000 to $10000 is shown for a better depiction. In this case, mean (1422) is higher than median (444) and plor looks right-skewed.

## Univariate Analysis of some catagorical variables

## *subscription status*

```{r, echo=FALSE, message=FALSE, warning=FALSE}
bank %>%
  group_by(y) %>%
  count() %>%
  mutate(percentage=n/nrow(bank)*100)
```


This univariate analysis could give a clue about probability of subscription in customers. from the above table, it is obvious that magority of customers did not subscribed (88.476%) comparing slighly more than ten percent of customers who finally choose to subscribe (11.524%). The good point about dataset is it provided valid outcome for all customers (whether subscribed or not).



## Bivariate Analysis

In this part two variables are considering and their relationship are examined. A dummy variable has been defined for outcome so that

0=No (did not subscribed), and

1=yes (subscribed).


## Corrolation between age and balance

```{r, echo=FALSE, message=FALSE, warning=FALSE}

cor_age_balance <- cor(bank$age, bank$balance)

paste("Corrolation between age and balance: ", cor_age_balance)

```

It shows small positive corrolation between age and balance. generally older clients tends to have higher balamce but the corrolation is more like a random relation  because it is more close to zero (random case) than 1 (strong corrolation).

## *subscription status~balance*

```{r, echo=FALSE, message=FALSE, warning=FALSE}

bank %>%
  filter(between(balance,-4000,30000)) %>%
  ggplot(aes(balance, result, color=y))+
  geom_point()

```

In this plot, subscription status versus balance depicted. Very high balances (higher than $30000) considered as outlier data and they have been omitted from plotting. Considering this plot alone is not very informative for me. I barely could decide if higher balance increase/decrease the chance of subscription by looking at this plot. For this reason, a Generalized linear line added to plot for a better information.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
bank %>%
  filter(between(balance,-4000,30000)) %>%
  ggplot(aes(balance, result))+
  geom_point()+
  geom_smooth(method = "glm", se=FALSE)
```

The added layer to this plot, comparing the previous plot, shows that those who has higher balance are more likely to subscribe to a term deposit. But how strong is this relationship? In other word, is balance very determinant parameter? Let's perform generalized analysis and see the parameters.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
mdl_bank <- glm(result ~ balance , data=bank, family = binomial)

mdl_bank %>%
  summary()
```

## *subscription status~age*

```{r, echo=FALSE, message=FALSE, warning=FALSE}

bank %>%
  filter(between(balance,-4000,30000)) %>%
  ggplot(aes(balance, result, color=y))+
  geom_point()#+
  #geom_smooth(method = "glm", se=FALSE)

```

## Generalized Linear Model





## *subscription status~age*


```{r, echo=FALSE, message=FALSE, warning=FALSE}

bank %>%
  ggplot(aes(age, result))+
  geom_point()+
  geom_smooth(method = "glm", se=FALSE)

```






```{r, echo=FALSE, message=FALSE, warning=FALSE}


bank2 <- read_csv("bank.csv") %>%
  mutate(result=ifelse(y == 'yes', 1, 0)) %>%
  filter (contact == "cellular" | result == "1" )  %>%
  filter (campaign == "1"       | result == "1" ) %>%
  filter (campaign == "1"       | result == "1" ) %>%
  filter (previous != "1"       | result == "1" ) %>%
  filter (previous != "2"       | result == "1" ) %>%
  filter (previous != "3"       | result == "1" ) %>%
  filter (previous != "4"       | result == "1" ) %>%
  filter (previous != "5"       | result == "1" ) #%>%
  #group_by(result)  %>%
  #count()
  





```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
bank2 %>%
  filter(between(balance,-4000,10000)) %>%
  ggplot(aes(balance, result))+
  geom_point()+
  geom_smooth(method = "glm", se=FALSE)
```




```{r, echo=FALSE, message=FALSE, warning=FALSE}
bank2 %>%
  
  ggplot(aes(age, result))+
  geom_point()+
  geom_smooth(method = "glm", se=FALSE)
```


