---
title: 'Assignment A02: Classification Models'
author: Ramin Ala
date: '2021-07-15'
slug: []
categories: []
tags:
  - Assignment
  - Classification
---
***Ramin Ala***


```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)

bank <- read_csv("bank.csv") %>%
  select(c(age, job, marital, education, default, balance, housing, loan, y)) %>%
  mutate(subscription_status=ifelse(y == 'yes', 1, 0))
 

```

## Executive Summary

The data is related with direct marketing campaigns (based on phone calls) of a Portuguese banking institution. This really is a **supervised question** because it has an outcome (y), which defined as if the product (bank term deposit) would be (or not) subscribed, "yes" means subscribed, and "no" means did not subscribe. This dataset consists of **`r nrow(bank)` observations**.

In addition to the outcome classification variable, there are 16 variables. Eight variables are related to statistical information about current and previous campaigns, and they seems irrelevant to  dependent variable. So, there are eight determinant variables Including:

Customer's ***age*** and ***balance*** (numerical variables),

Customer's ***job***, ***marital*** status, ***education***  (categorical variables)
 
***housing loan***, ***personal loan***, and ***default status*** (binary variables)  

Information regarding the reference of dataset is available at:

http://www3.dsi.uminho.pt/pcortez/bib/2011-esm-1.txt

## Univariate Analysis of numerical variables

## *age*

```{r, echo=FALSE, message=FALSE, warning=FALSE}
bank %>%
  ggplot(aes(age)) +
  geom_density(fill = "green", alpha = 0.4, size=0.01) +
   scale_x_continuous(breaks = seq(15, 90, by = 5))

summary(bank$age) %>%
  tidy() 
```

Above plot and table give an impression on the age of examined group. The minimum, maximum and average ages are 19, 87, and 41 years respectively. Majority of observations are concentrated at the age of between 30 to 35 years. This information would not provide any insight on the possible outcome.


## *balance*

```{r, echo=FALSE, message=FALSE, warning=FALSE}
bank %>%
  filter(between(balance,-4000,10000)) %>%
  ggplot(aes(balance)) +
  geom_histogram(fill = "red", alpha = 0.4, size=1) +
   scale_x_continuous(breaks = seq(-4000, 10000, by = 1000))

summary(bank$balance) %>%
  tidy()
```

Balance is generally between 0 to $500, however, amounts as high as $71000 is available in data. In the above plot, Only range from $-4000 to $10000 is shown for a better depiction. In this case, mean (1422) is higher than median (444) and plor looks right-skewed.

## Univariate Analysis of some catagorical variables

## *subscription status*

```{r, echo=FALSE, message=FALSE, warning=FALSE}
bank %>%
  group_by(y) %>%
  count() %>%
  mutate(percentage=n/nrow(bank)*100)
```


This univariate analysis could give a clue about probability of subscription in customers. from the above table, it is obvious that majority of customers did not subscribed (88.476%) comparing slightly more than ten percent of customers who finally choose to subscribe (11.524%). The good point about dataset is it provided valid outcome for all customers (whether subscribed or not).



## Bivariate Analysis

In this part two variables are considering and their relationship are examined. A dummy variable has been defined for outcome so that

0=No (did not subscribed), and

1=yes (subscribed).


## correlation between age and balance

```{r, echo=FALSE, message=FALSE, warning=FALSE}

cor_age_balance <- cor(bank$age, bank$balance)

paste("correlation between age and balance: ", cor_age_balance)

```

It shows small positive correlation between age and balance. generally older clients tends to have higher balance but the correlation is more like a random relation  because it is more close to zero (random case) than 1 (strong correlation).



## *subscription status~job*

```{r, echo=FALSE, message=FALSE, warning=FALSE}
bank0 <- bank %>%
  group_by(job, subscription_status) %>%
  count() 

bank0 %>%
  
  ggplot(aes(n, job,fill=as.factor(subscription_status)))+
  geom_col(stat = "identity")+
  
  geom_label(aes(label = n),
               position = position_stack(vjust = 0.5),
               size = 3.75,
               colour = 'black')

```

There are twelve categories of jobs.Managements, blue-collars, and technicians are the most frequent jobs among respondents. As shown in the plot, among these jobs, managements are more likely to subscribe. However, relative number of subscription is very high among students, unknown and unemployed people. These groups are more likely to subscribe. 

## *subscription status~marital*


```{r, echo=FALSE, message=FALSE, warning=FALSE}
bank0 <- bank %>%
  group_by(subscription_status, marital) %>%
  count()  %>% 
mutate(percent = n)
  


bank0 %>%
  
  ggplot(aes(marital, n,fill=as.factor(subscription_status)))+
  geom_col(stat = "identity")+
  
  geom_label(aes(label = n),
               position = position_stack(vjust = 0.5),
               size = 3.75,
               colour = 'black')

```

Similar to previous analysis could be done for marital status.This plot shows majority of respondents are married. However, divorced group are more likely to subscribe when numbers compared in each group.




## *subscription status~balance*

```{r, echo=FALSE, message=FALSE, warning=FALSE}

bank %>%
  filter(between(balance,-4000,30000)) %>%
  ggplot(aes(balance, subscription_status, color=y))+
  geom_point()

```

In this plot, subscription status versus balance depicted. Very high balances (higher than $30000) considered as outlier data and they have been omitted from plotting. Considering this plot alone is not very informative for me. I barely could decide if higher balance increase/decrease the chance of subscription by looking at this plot. For this reason, a Generalized linear line added to plot for a better information.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
bank %>%
  filter(between(balance,-4000,30000)) %>%
  ggplot(aes(balance, subscription_status))+
  geom_point()+
  geom_smooth(method = "glm", se=FALSE)
```

The added layer to this plot, comparing the previous plot, shows that those who has higher balance are more likely to subscribe to a term deposit. But how strong is this relationship? In other word, is balance very determinant parameter? Let's perform generalized analysis and see the parameters.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
mdl_bank <- glm(subscription_status ~ balance , data=bank, family = binomial)

mdl_bank %>%
  summary()
```



## *subscription status~age*

```{r, echo=FALSE, message=FALSE, warning=FALSE}

bank %>%
  ggplot(aes(age, subscription_status))+
  geom_point()+
  geom_smooth(method = "glm", se=FALSE)

```

Again it looks difficult to get information from this plot as they have not strong relation. However, blue line that is the generalized regression line, shows that higher age people tends to subscribe more. An important subject to remember is in a multi-variable problem, considering one parameter alone could be misleading because of correlation of variables.

## *subscription status~default*



```{r, echo=FALSE, message=FALSE, warning=FALSE}
ss <- bank %>%
  group_by(default) %>%
  count(subscription_status) 


 ss$default_result <-paste(ss$default,ss$subscription_status)

ss %>%
  ggplot(aes(default_result, n))+
  geom_col(fill="blue")


```


In this plot combination of two categorical variables are counted. One is default and the other is dependent variable (result). We have four categories here (four possible combinations for two binary variables). It seems that default clients (default=yes) is very unlikely to subscribe.

## Analysis of three variables: result versus age and balance


```{r, echo=FALSE, message=FALSE, warning=FALSE}
bank %>%
  filter(between(balance,-4000,30000)) %>%
  ggplot(aes(age, balance, color=y))+
  geom_point()
```
This plot consists of two numerical variables (age and balance) that both are explanatory variables and response variable which depicted by color (pink colors did not subscribed and blue colors subscribed). 


## *Analysis of four variables:subscription status~age and marital and loan*

```{r, echo=FALSE, message=FALSE, warning=FALSE}

bank %>%
  ggplot(aes(age, subscription_status, color=marital, shape=loan))+
  geom_point()

```

This plot probably give not more information than a two variable analysis because point are very closed hard to distinguish. So it time to examine all parameters using glm function.



## Generalized Linear Model


```{r, echo=FALSE, message=FALSE, warning=FALSE}
mdl_bank <- glm(subscription_status ~ age+job+marital+education+default+balance+housing+loan , data=bank, family = binomial)

mdl_bank %>%
  summary()
```

Pr(>|z|) value for  balance is close to 1 and this makes us to believe it is not a determinant parameter, but this parameter is negligible for housing with relatively high value for its coefficient. This is the reason for importance of housing status. A similar circumstance is for loan status.

In addition, Coefficients of above analysis are all small because the number of respondents with subscribed status are very lower than not subscribed clients.

Finally, some parameters that seemed to be determinant in univariate analysis are not important in fact. This is because of correlation between non determinant and determinant 
variables.

## Summary of findings

- Most respondents in this campaign was 30 to 35 years old with near zero balance.Besides, nearly 90% of respondents did not subscribe finally.

- Univariate analysis shows positive slope for both age and balance (higher age/balance means higher chance of subscription). However, generalized linear Modeling with consideration of all variables shows balance has not an impact on chance of subscription.

- Univariate analysis  of job shows that a big portion of students, unemployed, and those with unknown jobs subscribed, but respondents with blue-color jobs and technicians did not subscribe so often.

- Divorced respondents was more likely to subscribe than married and single ones.

- Generalized Linear Model shows that balance is not a determinant factor but housing and loan status are both determinant.

- Coefficients of glm analysis are all small because the number of respondents with subscribed status are very lower than not subscribed clients.







































## Resampling Methods

```{r, warning=FALSE, message=FALSE}
library(caret)

library(yardstick)

```

## Cross-Validation method

## LOOCV (Leave one out cross validation) 

The LOOCV method is using all data points therefore it reduces the bias.

However, the execution time could be so long if n (number of observation) is large. Besides, we know that low bias means high variance and the model is more prone to overfitting.

---

```{r, warning=FALSE}
# Define training control
train.control_LOOCV <- trainControl(method = "LOOCV")

# Train the model
model_LOOCV <- train(subscription_status ~ ., data = bank, method = "lm",
               trControl = train.control_LOOCV)
# Summarize the results
print(model_LOOCV)
```


## K-fold cross-validation


```{r, warning=FALSE}
# Define training control
set.seed(120) 

train.control_K_fold <- trainControl(method = "repeatedcv", number = 10, repeats = 4)

# Train the model
model_K_fold <- train(subscription_status ~., data = bank, method = "lm",
               trControl = train.control_K_fold)
# Summarize the results
print(model_K_fold)
```



## Bootstrap method 


```{r, warning=FALSE}
# Define training control
train.control_Bootstrap <- trainControl(method = "boot", number = 100)

# Train the model
model_Bootstrap <- train(subscription_status ~., data = bank, method = "lm",
               trControl = train.control_Bootstrap)
# Summarize the results
print(model_Bootstrap)
```



## Evaluating performance with yardstick


```{r}

status_predictions_SLR <- predict(SLR, new_data = bank)
status_predictions_MLR <- predict(MLR, new_data = bank)
status_predictions_model_LOOCV <- predict(model_LOOCV$finalModel, new_data = bank)
status_predictions_model_K_fold <- predict(model_K_fold$finalModel, new_data = bank)
status_predictions_model_Bootstrap <- predict(model_Bootstrap$finalModel, new_data = bank)



test_results_SLR <- bank %>%
  bind_cols(status_pred=status_predictions_SLR)

test_results_MLR <- bank %>%
  bind_cols(status_pred=status_predictions_MLR)

test_results_model_LOOCV <- bank %>%
  bind_cols(status_pred=status_predictions_model_LOOCV)

test_results_model_K_fold <- bank %>%
  bind_cols(status_pred=status_predictions_model_K_fold)

test_results_model_Bootstrap <- bank %>%
  bind_cols(status_pred=status_predictions_model_Bootstrap)



# Calculate the confusion matrix
conf_mat(telecom_results, truth = subscription_status,
    estimate = .pred_class)
```

