---
title: 'Assignment A01: Regression Models'
author: Ramin Ala
date: '2021-07-08'
slug: []
categories: []
tags:
  - Regression
  - Assignment
---

```{r}
library(tidyverse)
library(readxl)
library(tidymodels)
#diamonds_price <- read_excel("MBA6636_SM21_Professor_Proposes_Data.xlsx", col_names = TRUE, col_types = NULL, na = "", skip = 0)

diamonds <- read_excel("MBA6636_SM21_Professor_Proposes_Data.xlsx") %>%
  janitor::clean_names() %>%
  subset(select = -wholesaler)

names(diamonds)

```

```{r}
diamonds %>%
  ggplot(aes(carat, price))+
  geom_point()+
  geom_smooth(method="lm", se= FALSE)
```

```{r}
diamonds %>%
  group_by(colour) %>%
  summarise(price_mean=mean(price)) %>%
  ggplot(aes(colour,price_mean))+
  geom_col(fill="blue")
```

```{r}
yy <- diamonds %>%
  group_by(clarity) %>%
  summarise(price_mean=mean(price)) %>%
  arrange(desc(price_mean))
yy
yy %>%
  ggplot(aes(factor(clarity),price_mean))+
  geom_col(fill="yellow")
```

```{r}
diamonds %>%
  group_by(cut) %>%
  summarise(price_mean=mean(price)) %>%
  ggplot(aes(cut,price_mean))+
  geom_col(fill="red")
```





## Summary of Findings:

## Identify Determinants of Price


## Comments on distribution of data


## Univariate Analysis

## Metric Data: Price, Carat


## Non-Metric Data: Color, Cut, Certification, Polish, Symmetry, Clarity


## Bivariate Analysis:

## Correlation, Covariance

These are measures of assosiation.

## Regression Model

## Simple and Multiple Linear Regression

A linear model is generated using **lm()** function. Seven variables passes to model and intercept defined as zero. As result, coefficients calculated 

```{r}
MLR <- lm(price~ .+0, data=diamonds)
summary(MLR)  %>%
  tidy()

```


 "carat"  "colour"  "clarity"    "cut"   "certification" "polish"  "symmetry" 

```{r}

explanatory_data= data.frame(carat= 0.9,
                             colour= "J",
                             clarity= "SI2",
                             cut= "V",
                             certification= "GIA",
                             polish= "G",
                             symmetry= "V")
  

predict(MLR, explanatory_data)
```


