---
title: Communicating with Data in the Tidyverse
author: ''
date: '2021-09-09'
slug: []
categories: []
tags:
  - Markdown
---

```{r results='hide'}
library(tidyverse)

ilo_hourly_compensation <- read.csv("ilo_hourly_compensation.csv")
ilo_working_hours <- read.csv("ilo_working_hours.csv")
```

## 1-2 Join the two data sets together

In the video, you have learned that the inner_join() function of dplyr needs to be given a "key" on which two data frames are joined. Actually, multiple keys that need to match can be specified. In this first exercise, you are going to join two data sets by two keys. The data frames ilo_hourly_compensation and ilo_working_hours are already loaded for you and are available in your workspace.

This course touches on a lot of concepts you may have forgotten, so if you ever need a quick refresher, download the tidyverse Cheat Sheet and keep it handy!


```{r results='hide'}
# Join both data frames
ilo_data <- ilo_hourly_compensation %>%
  inner_join(ilo_working_hours, by = c("country", "year"))

# Count the resulting rows
ilo_data  %>% 
    count()

# Examine ilo_data
ilo_data
```

## 1-3 Change variable types

For displaying data in ggplot2 graphics, it is often helpful to convert all the variables to the right data type. Usually, categorical variables like country in this example should be converted to factors before plotting them. You can do so using as.factor(). In your data set, two columns are still of type "character" – use mutate() to turn them into factors.

```{r}
# Turn year and country into a factor
ilo_data_corrected <- ilo_data %>%
  mutate(year = as.factor(as.numeric(year)),
        country = as.factor(country))

# See the results
ilo_data_corrected
```

## 1-5 Filter the data for plotting

Use the filter() function of dplyr to remove non-European countries. A vector called european_countries has been created that contains all European countries which should remain in the data set. Here, the %in% operator introduced in the video might come in handy. Note that only a subset of European countries for which enough data exist is retained. For instance, Poland is missing.

```{r}

```

## 1-6 Some summary statistics

Use dplyrs group_by() and summarize() to compute summary statistics for both years.

```{r}

```

## 1-7 A basic scatter plot

In this exercise, you will create a very basic scatter plot with ggplot2. This is mostly a repetition of stuff you've learnt in the prerequisites for this course, so it should be easy for you.

```{r}

```

## 1-8 Add labels to the plot

As mentioned in the video, you're going to enhance the plot from the previous exercise by adding a title, a subtitle, and a caption to the plot as well as giving the axes meaningful names. You're going to use the labs() function to do this – try using ?labs in the console to see the available options.

```{r}

```

## 1-10 Apply a default theme

As you've learnt in the videos, ggplot2 comes with a set of predefined themes. Try out some of them!

```{r}

```

## 1-11 Change the appearance of titles

Besides applying defined theme presets, you can tweak your plot even more by altering different style attributes of it. Hint: You can reuse and overwrite the ilo_plot variable generated in the previous exercise – the current plot is already shown in the window on the right.

```{r}

```

## 1-12 Alter background color and add margins

Further customize the look of your plot with more arguments to the theme function call. You can continue working on your ilo_plot object created in the last exercise.

```{r}

```

