---
title: Introduction to Regression in R
author: ''
date: '2021-07-07'
slug: []
categories: []
tags:
  - Regression
---

```{r}
library(tidyverse)
library(broom)
```



```{r}
taiwan_real_estate <- read_csv("taiwan_real_estate.csv")
sp500_yearly_returns <- read_csv("sp500_yearly_returns.csv")

ad_conversion <- read_csv("ad_conversion.csv")

churn <- read_csv("churn.csv")

```

# 1-3 Visualizing two variables

points 50% transparent by setting alpha to 0.5.
adding a trend line, calculated using a linear regression.

```{r}
taiwan_real_estate %>%
ggplot(aes(n_convenience, price_twd_msq))+
geom_point(alpha=0.5)+
  geom_smooth(method="lm", se=FALSE)
```

# 1-7 Linear regression with lm()


```{r}
# Run a linear regression of price_twd_msq vs. n_convenience
lm(price_twd_msq~n_convenience, taiwan_real_estate)
```


# 1-9 Visualizing numeric vs. categorical


```{r}
# Using taiwan_real_estate, plot price_twd_msq
taiwan_real_estate %>% ggplot(aes(price_twd_msq)) +
  # Make it a histogram with 10 bins
  geom_histogram(bins=10) +
  # Facet the plot so each house age group gets its own panel
  facet_wrap(vars(house_age_years))
```

# 1-10 Calculating means by category


```{r}
summary_stats <- taiwan_real_estate %>% 
  # Group by house age
  group_by(house_age_years) %>% 
  # Summarize to calculate the mean house price/area
  summarise(mean_by_group = mean(price_twd_msq))

# See the result
summary_stats
```

# 1-11 lm() with a categorical explanatory variable


```{r}
# Run a linear regression of price_twd_msq vs. house_age_years
mdl_price_vs_age <- lm(price_twd_msq~house_age_years, taiwan_real_estate)

# See the result
mdl_price_vs_age


#---------------------------------

# Update the model formula to remove the intercept
mdl_price_vs_age_no_intercept <- lm(
  price_twd_msq ~ house_age_years+0, 
  data = taiwan_real_estate
)

# See the result
mdl_price_vs_age_no_intercept
```

# 2-2 Predicting house prices

you specify values for each of the explanatory variables, feed them to the model, and you get a prediction for the corresponding response variable. The code flow is as follows:

explanatory_data <- tibble(
  explanatory_var = some_values
)
explanatory_data %>%
  mutate(
    response_var = predict(model, explanatory_data)
  )
  
  
```{r}
taiwan_real_estate


mdl_price_vs_conv <- lm(formula = price_twd_msq ~ n_convenience, data = taiwan_real_estate)



# Create a tibble with n_convenience column from zero to ten
explanatory_data <- tibble(n_convenience = 0:10)
explanatory_data
# Use mdl_price_vs_conv to predict with explanatory_data
prediction_data <- explanatory_data %>% 
  mutate(
    price_twd_msq = predict(mdl_price_vs_conv, explanatory_data)
  )

# See the result
prediction_data
```
  
# 2-3 Visualizing predictions
  
  
```{r}
# Add to the plot
ggplot(taiwan_real_estate, aes(n_convenience, price_twd_msq)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  # Add a point layer of prediction data, colored yellow
  geom_point(data = prediction_data,
color = "yellow")
```
  
## 2-6 Extracting model elements


```{r, echo=FALSE, message=FALSE, results='hide'}
# Get the model coefficients of mdl_price_vs_conv
coefficients(mdl_price_vs_conv)

# Get the fitted values of mdl_price_vs_conv
fitted(mdl_price_vs_conv)

# Get the residuals of mdl_price_vs_conv
residuals(mdl_price_vs_conv)

# Print a summary of mdl_price_vs_conv
summary (mdl_price_vs_conv)
```

## 2-7 Manually predicting house prices

```{r}
# Get the coefficients of mdl_price_vs_conv
coeffs <- coefficients(mdl_price_vs_conv)

# Get the intercept
intercept <- coeffs[1]

# Get the slope
slope <- coeffs[2]

explanatory_data %>% 
  mutate(
    # Manually calculate the predictions
    price_twd_msq = slope*n_convenience+intercept
  )

# Compare to the results from predict()
predict(mdl_price_vs_conv, explanatory_data)
```

## 2-8 Using broom

Many programming tasks are easier if you keep all your data inside data frames. This is particularly true if you are a tidyverse fan, where dplyr and ggplot2 require you to use data frames. The broom package contains functions that decompose models into three data frames: one for the coefficient-level elements (the coefficients themselves, as well as p-values for each coefficient), the observation-level elements (like fitted values and residuals), and the model-level elements (mostly performance metrics).

The functions in broom are generic. That is, they work with many model types, not just linear regression model objects. They also work with logistic regression model objects (as you'll see in Chapter 4), and many other types of model.


```{r}
# Get the coefficient-level elements of the model
tidy(mdl_price_vs_conv)

# Get the observation-level elements of the model
augment(mdl_price_vs_conv)

# Get the model-level elements of the model
glance(mdl_price_vs_conv)
```






## 2-11 Plotting consecutive portfolio returns

Regression to the mean is also an important concept in investing. Here you'll look at the annual returns from investing in companies in the Standard and Poor 500 index (S&P 500), in 2018 and 2019.

```{r}
# Using sp500_yearly_returns, plot return_2019 vs. return_2018
sp500_yearly_returns %>% ggplot(aes(return_2018, return_2019)) +
  # Make it a scatter plot
  geom_point() +
  # Add a line at y = x, colored green, size 1
  geom_abline(color = "green", size = 1) +
  # Add a linear regression trend line, no std. error ribbon
  geom_smooth(method="lm", se=FALSE) +
  # Fix the coordinate ratio
  coord_fixed()
```

## 2-12 Modeling consecutive returns

Let's quantify the relationship between returns in 2019 and 2018 by running a linear regression and making predictions. By looking at companies with extremely high or extremely low returns in 2018, we can see if their performance was similar in 2019.

```{r}
# Run a linear regression on return_2019 vs. return_2018
# using sp500_yearly_returns
mdl_returns <- lm(return_2019~return_2018, sp500_yearly_returns)

# See the result
mdl_returns

#-----------------------------------------
# Create a data frame with return_2018 at -1, 0, and 1 
explanatory_data <- tibble(return_2018=-1:1)



# Use mdl_returns to predict with explanatory_data
predict(mdl_returns, explanatory_data)


```

## 2-14 Transforming the explanatory variable

If there is no straight line relationship between the response variable and the explanatory variable, it is sometimes possible to create one by transforming one or both of the variables. Here, you'll look at transforming the explanatory variable.

You'll take another look at the Taiwan real estate dataset, this time using the distance to the nearest MRT (metro) station as the explanatory variable. You'll use code to make every commuter's dream come true: shortening the distance to the metro station by taking the square root. Take that, geography!


```{r}
# Run the code to see the plot
# Edit so x-axis is square root of dist_to_mrt_m
ggplot(taiwan_real_estate, aes(dist_to_mrt_m, price_twd_msq)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```


edit code

```{r}
# Run the code to see the plot
# Edit so x-axis is square root of dist_to_mrt_m
ggplot(taiwan_real_estate, aes(sqrt(dist_to_mrt_m), price_twd_msq)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
# Run a linear regression of price_twd_msq vs. 
# square root of dist_to_mrt_m using taiwan_real_estate
mdl_price_vs_dist <- lm(price_twd_msq~sqrt(dist_to_mrt_m), taiwan_real_estate)

# See the result
mdl_price_vs_dist
```
```{r}
# From previous step
mdl_price_vs_dist <- lm(
  price_twd_msq ~ sqrt(dist_to_mrt_m), 
  data = taiwan_real_estate
)

# Use this explanatory data
explanatory_data <- tibble(
  dist_to_mrt_m = seq(0, 80, 10) ^ 2
)

# Use mdl_price_vs_dist to predict explanatory_data
prediction_data <- explanatory_data %>%
mutate(
  price_twd_msq= predict(mdl_price_vs_dist,explanatory_data)



)
# See the result
prediction_data


ggplot(taiwan_real_estate, aes(sqrt(dist_to_mrt_m), price_twd_msq)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  # Add points from prediction_data, colored green, size 5
  geom_point(data=prediction_data, color = "green", size= 5)


```


## 2-15 ransforming the response variable too

The response variable can be transformed too, but this means you need an extra step at the end to undo that transformation. That is, you "back transform" the predictions.

In the video, you saw the first step of the digital advertising workflow: spending money to buy ads, and counting how many people see them (the "impressions"). The next step is determining how many people click on the advert after seeing it.

```{r}
# Run the code to see the plot
# Edit to raise x, y aesthetics to power 0.25
ggplot(ad_conversion, aes(n_impressions, n_clicks)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
# Run the code to see the plot
# Edit to raise x, y aesthetics to power 0.25
ggplot(ad_conversion, aes(n_impressions^0.25, n_clicks^0.25)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
# From previous steps
mdl_click_vs_impression <- lm(
  I(n_clicks ^ 0.25) ~ I(n_impressions ^ 0.25),
  data = ad_conversion
)
explanatory_data <- tibble(
  n_impressions = seq(0, 3e6, 5e5)
)
prediction_data <- explanatory_data %>% 
  mutate(
    n_clicks_025 = predict(mdl_click_vs_impression, explanatory_data),
    n_clicks = n_clicks_025 ^ 4
  )

ggplot(ad_conversion, aes(n_impressions ^ 0.25, n_clicks ^ 0.25)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  # Add points from prediction_data, colored green
  geom_point(data=prediction_data,  color="green")
```

## 3-2 Coefficient of determination

The coefficient of determination is a measure of how well the linear regression line fits the observed values. For simple linear regression, it is equal to the square of the correlation between the explanatory and response variables.

```{r}
# Print a summary of mdl_click_vs_impression
summary(mdl_click_vs_impression)

#--------------------

# Get coeff of determination for mdl_click_vs_impression_orig
mdl_click_vs_impression %>% 
  # Get the model-level details
  glance() %>%
 
  # Pull out r.squared
  pull(r.squared)

```

## 3-3 Residual standard error

Residual standard error (RSE) is a measure of the typical size of the residuals. Equivalently, it's a measure of how badly wrong you can expect predictions to be. Smaller numbers are better, with zero being a perfect fit to the data.

```{r}
# Get RSE for mdl_click_vs_impression
mdl_click_vs_impression %>% 
  # Get the model-level details
  glance() %>% 
  # Pull out sigma
  pull(sigma)
```

## 3-8 Drawing diagnostic plots

It's time for you to draw these diagnostic plots yourself. Let's go back to the Taiwan real estate dataset and the model of house prices versus number of convenience stores.

Recall that autoplot() lets you specify which diagnostic plots you are interested in.

1 residuals vs. fitted values
2 Q-Q plot
3 scale-location
mdl_price_vs_conv is available, and ggplot2 and ggfortify are loaded.


## 4-2 Exploring the explanatory variables

When the response variable is logical, all the points lie on the y equals zero and y equals one lines, making it difficult to see what is happening. In the video, until you saw the trend line, it wasn't clear how the explanatory variable was distributed on each line. This can be solved with a histogram of the explanatory variable, faceted on the response.

```{r}
# Using churn, plot time_since_last_purchase
ggplot(churn, aes(time_since_last_purchase)) +
  # as a histogram with binwidth 0.25
  geom_histogram(binwidth=0.25) +
  # faceted in a grid with has_churned on each row
  facet_grid(vars(has_churned))


# Redraw the plot with time_since_first_purchase

ggplot(churn, aes(time_since_first_purchase)) +
  geom_histogram(binwidth=0.25) +
   facet_grid(vars(has_churned))
```



## 4-3 Visualizing linear and logistic models

As with linear regressions, ggplot2 will draw model predictions for a logistic regression without you having to worry about the modeling code yourself. To see how the predictions differ for linear and logistic regressions, try drawing both trend lines side by side. Spoiler: you should see a linear (straight line) trend from the linear model, and a logistic (S-shaped) trend from the logistic model.

```{r}
# Using churn plot has_churned vs. time_since_first_purchase
churn %>% ggplot(aes(time_since_first_purchase, has_churned)) +
  # Make it a scatter plot
  geom_point() +
  # Add an lm trend line, no std error ribbon, colored red
  geom_smooth(method="lm", se=FALSE, color="red")+

# Add a glm trend line, no std error ribbon, binomial family
  geom_smooth(method = "glm", se = FALSE,method.args = list(family = binomial))

```


## 4-4 Logistic regression with glm()

Linear regression and logistic regression are special cases of a broader type of models called generalized linear models ("GLMs"). A linear regression makes the assumption that the residuals follow a Gaussian (normal) distribution. By contrast, a logistic regression assumes that residuals follow a binomial distribution.

Here, you'll model how the length of relationship with a customer affects churn.

```{r}
# Fit a logistic regression of churn vs. 
# length of relationship using the churn dataset
mdl_churn_vs_relationship <- glm(has_churned~time_since_first_purchase, data=churn, family = binomial)

# See the result
mdl_churn_vs_relationship
```


## 4-6 Probabilities

There are four main ways of expressing the prediction from a logistic regression model – we'll look at each of them over the next four exercises. Firstly, since the response variable is either "yes" or "no", you can make a prediction of the probability of a "yes". Here, you'll calculate and visualize these probabilities.

```{r}
# Make a data frame of predicted probabilities

#prediction_data <- explanatory_data %>% mutate(has_churned = predict(mdl_churn_vs_relationship, explanatory_data, type = "response"))

# See the result
#prediction_data
```



