---
title: 'Case Study: Exploratory Data Analysis in R'
author: ''
date: '2021-08-28'
slug: []
categories: []
tags:
  - data_visualization
  - data_wrangling
---


```{r}
library(tidyverse)
votes <- read.csv("votes.csv")
```

## 1-2 Filtering rows

The vote column in the dataset has a number that represents that country's vote:

1 = Yes
2 = Abstain
3 = No
8 = Not present
9 = Not a member
One step of data cleaning is removing observations (rows) that you're not interested in. In this case, you want to remove "Not present" and "Not a member".

```{r}
# Filter for votes that are "yes", "abstain", or "no"
votes %>% filter(vote<=3)
```

## 1-3 Adding a year column

The next step of data cleaning is manipulating your variables (columns) to make them more informative.

In this case, you have a session column that is hard to interpret intuitively. But since the UN started voting in 1946, and holds one session per year, you can get the year of a UN resolution by adding 1945 to the session number.

```{r}
# Add another %>% step to add a year column
votes %>%
  filter(vote <= 3) %>%
mutate(year=session+1945)
```

## 1-4 Adding a country column

The country codes in the ccode column are what's called Correlates of War codes. This isn't ideal for an analysis, since you'd like to work with recognizable country names.

You can use the countrycode package to translate. For example:

library(countrycode)

# Translate the country code 2
> countrycode(2, "cown", "country.name")
[1] "United States"

# Translate multiple country codes
> countrycode(c(2, 20, 40), "cown", "country.name")
[1] "United States" "Canada"        "Cuba"

```{r}
# Load the countrycode package
library(countrycode)

# Convert country code 100
countrycode(100, "cown", "country.name")

# Add a country column within the mutate: votes_processed
votes_processed <- votes %>%
  filter(vote <= 3) %>%
  mutate(year = session + 1945,
         country = countrycode(ccode, "cown", "country.name"))
```

