---
title: Project
author: ''
date: '2021-06-22'
slug: []
categories: []
tags:
  - data_visualization
  - data_wrangling
  - R
---

**Ramin Ala**





```{r packages, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(tidymodels) #Interesting collections of models.
library(ggtext) #similar to ggplot but for text.
library(knitr)
library(kableExtra) 
library(janitor)
library(gganimate)
library(gifski)
```



---

## Import Data: Air Quality


```{r, echo=FALSE, message=FALSE, include=TRUE, warning=FALSE}
air <- read_csv("AirQualityUCI.csv", na = c("n/a", "", "NA")) %>%
  janitor::clean_names() 
names(air)
```
This post consists of a bit of data clean up and exploratory analysis of a dataset available in University of californium, Irvine's website. The dataset includes `r ncol(air)` variables and `r nrow(air)` observations that represent some parameters related to air quality. This dataset is selected to perform a basic fitting.

This dataset could be find in the below link:

https://archive.ics.uci.edu/ml/datasets/Air+Quality

```{r}
ee <- air %>%
  select(co_gt,pt08_s1_co) %>%
subset(co_gt!=-200 & pt08_s1_co!=-200) %>%
 
  ggplot(aes(co_gt,pt08_s1_co))+
  geom_point()+
  labs(x="C",y="T")

ee
```


## Goal: Predict T (tin oxide content) from C (co content)


$$\widehat{T}_{i} = \beta_0 + \beta_1 \times C_{i}$$

## fff

```{r}
ee <- air %>%
  select(co_gt,pt08_s1_co) %>%
subset(co_gt!=-200 & pt08_s1_co!=-200) %>%
 
  ggplot(aes(co_gt,pt08_s1_co))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "#8E2C90")+
  labs(x="C",y="T")

ee
```

## Specify model and Set model fitting *engine*

```{r}
linear_reg()

linear_reg() %>%
  set_engine("lm") # lm: linear model
```

## Fit model & estimate parameters

... using **formula syntax**

```{r fit-model}
linear_reg() %>%
  set_engine("lm") %>%
  fit(pt08_s1_co~co_gt, data = air) #in: inches #we want to predict height, using width
```


$$\widehat{T}_{i} = \beta_0 + \beta_1 \times C_{i}$$


$$\widehat{T}_{i} = 1055.0067 + 0.1759 \times C_{i}$$





```{r}
fgfg <- linear_reg() %>%
  set_engine("lm") %>%
  fit(co_gt~pt08_s1_co, data = air) %>%
  tidy()


fgfg
```
