---
title: GGPLOT2
author: ''
date: '2021-06-13'
slug: ggplot2
categories:
  - ggplot2
tags: []
---


# 1-1 Drawing your first plot

To get a first feel for ggplot2, let's try to run some basic ggplot2 commands. The mtcars dataset contains information on 32 cars from a 1973 issue of Motor Trend magazine. This dataset is small, intuitive, and contains a variety of continuous and categorical variables.

```{r}
# Load the ggplot2 package
library(ggplot2)
# Explore the mtcars data frame with str()
str(mtcars)
names(mtcars)
# Execute the following command
ggplot(mtcars, aes(cyl, mpg)) +
  geom_point()
```

# 1-2 Data columns types affect plot types

The plot from the previous exercise wasn't really satisfying. Although cyl (the number of cylinders) is categorical, you probably noticed that it is classified as numeric in mtcars. This is really misleading because the representation in the plot doesn't match the actual data type. You'll have to explicitly tell ggplot2 that cyl is a categorical variable.


```{r}
# Load the ggplot2 package
library(ggplot2)

# Change the command below so that cyl is treated as factor
ggplot(mtcars, aes(factor(cyl), mpg)) +
  geom_point()
```


# 1-3 Mapping data columns to aesthetics

Let's dive a little deeper into the three main topics in this course: The data, aesthetics, and geom layers. We'll get to making pretty plots in the last chapter with the themes layer.

We'll continue working on the 32 cars in the mtcars data frame.

Consider how the examples and concepts we discuss throughout these courses apply to your own data-sets!

```{r}
# Change the color aesthetic to a size aesthetic
ggplot(mtcars, aes(wt, mpg, size = disp)) +
  geom_point()
```

# 1-4 Adding geometries

The diamonds dataset contains details of 1,000 diamonds. Among the variables included are carat (a measurement of the diamond's size) and price.

You'll use two common geom layer functions:

geom_point() adds points (as in a scatter plot).
geom_smooth() adds a smooth trend curve.
As you saw previously, these are added using the + operator.

ggplot(data, aes(x, y)) +
  geom_*()
Where * is the specific geometry needed.

```{r}
# Add geom_smooth() with +
ggplot(diamonds, aes(carat, price)) +
  geom_point() +
  geom_smooth()
```

# 1-5 Changing one geom or every geom
If you have multiple geoms, then mapping an aesthetic to data variable inside the call to ggplot() will change all the geoms. It is also possible to make changes to individual geoms by passing arguments to the geom_*() functions.

geom_point() has an alpha argument that controls the opacity of the points. A value of 1 (the default) means that the points are totally opaque; a value of 0 means the points are totally transparent (and therefore invisible). Values in between specify transparency.

The plot you drew last time is provided in the script.

```{r}
# Make the points 40% opaque
ggplot(diamonds, aes(carat, price, color = clarity)) +
  geom_point(alpha=0.4) +
  geom_smooth()
```

# 1-6 Saving plots as variables

Plots can be saved as variables, which can be added two later on using the + operator. This is really useful if you want to make multiple related plots from a common base.

```{r}
# From previous step
plt_price_vs_carat <- ggplot(diamonds, aes(carat, price))

# Edit this to map color to clarity
# Assign the updated plot to a new object
plt_price_vs_carat_by_clarity <- plt_price_vs_carat + geom_point(aes(color = clarity))

# See the plot
plt_price_vs_carat_by_clarity
```



## Aesthetics

# 2-1 All about aesthetics: color, shape and size

In the video you saw 9 visible aesthetics. Let's apply them to a categorical variable — the cylinders in mtcars, cyl.

These are the aesthetics you can consider within aes() in this chapter: x, y, color, fill, size, alpha, labels and shape.

One common convention is that you don't name the x and y arguments to aes(), since they almost always come first, but you do name other arguments.

In the following exercise the fcyl column is categorical. It is cyl transformed into a factor.

```{r}
ggplot(mtcars, aes(wt, mpg, color = cyl)) +
  # Set the shape and size of the points
  geom_point(shape=1, size=4)
```

# 2-2 All about aesthetics: color vs. fill

Typically, the color aesthetic changes the outline of a geom and the fill aesthetic changes the inside. geom_point() is an exception: you use color (not fill) for the point color. However, some shapes have special behavior.

The default geom_point() uses shape = 19: a solid circle. An alternative is shape = 21: a circle that allow you to use both fill for the inside and color for the outline. This is lets you to map two aesthetics to each point.

All shape values are described on the points() help page.

fcyl and fam are the cyl and am columns converted to factors, respectively.

```{r}
# Map color to am
ggplot(mtcars, aes(wt, mpg, fill = cyl,color=am)) +
  geom_point(shape = 21, size = 4, alpha = 0.6)

```

# 2-3 All about aesthetics: comparing aesthetics

Now that you've got some practice with using attributes, be careful of a major pitfall: these attributes can overwrite the aesthetics of your plot!

```{r}
# Base layer
plt_mpg_vs_wt <- ggplot(mtcars, aes(wt, mpg))

# Use text layer and map fcyl to label
plt_mpg_vs_wt +
  geom_text(aes(label = cyl))
```

# 2-4 All about attributes: color, shape, size and alpha

This time you'll use these arguments to set attributes of the plot, not map variables onto aesthetics.

You can specify colors in R using hex codes: a hash followed by two hexadecimal numbers each for red, green, and blue ("#RRGGBB"). Hexadecimal is base-16 counting. You have 0 to 9, and A representing 10 up to F representing 15. Pairs of hexadecimal numbers give you a range from 0 to 255. "#000000" is "black" (no color), "#FFFFFF" means "white", and `"#00FFFF" is cyan (mixed green and blue).

A hexadecimal color variable, my_blue has been defined for you.

```{r}
# A hexadecimal color
my_blue <- "#4ABEFF"

# Change the color mapping to a fill mapping
ggplot(mtcars, aes(wt, mpg, fill = cyl)) +
  # Set point size and shape
 geom_point(color=my_blue, size=10, shape=1)
```

# 2-5 All about attributes: conflicts with aesthetics

In the videos you saw that you can use all the aesthetics as attributes. Let's see how this works with the aesthetics you used in the previous exercises: x, y, color, fill, size, alpha, label and shape.

In this exercise you will set all kinds of attributes of the points!

You will continue to work with mtcars.

```{r}
ggplot(mtcars, aes(wt, mpg, color = fcyl)) +
  # Add points layer with shape 24 and color yellow
  geom_point(shape=24, color="yellow")
```

# 2-6 Going all out

In this exercise, you will gradually add more aesthetics layers to the plot. You're still working with the mtcars dataset, but this time you're using more features of the cars. Each of the columns is described on the mtcars help page. Columns fcyl and fam have been added (as before) as categorical versions of cyl and am respectively.

Notice that adding more aesthetic mappings to your plot is not always a good idea! You may just increase complexity and decrease readability.

```{r}
# 5 aesthetics: add a mapping of size to hp / wt
ggplot(mtcars, aes(mpg, qsec, color = cyl, size= hp/wt)) +
  geom_point()
```

# 2-7 Updating aesthetic labels

In this exercise, you'll modify some aesthetics to make a bar plot of the number of cylinders for cars with different types of transmission.

You'll also make use of some functions for improving the appearance of the plot.

labs() to set the x- and y-axis labels. It takes strings for each argument.
scale_color_manual() defines properties of the color scale (i.e. axis). The first argument sets the legend title. values is a named vector of colors to use.

```{r}
palette <- c(automatic = "#377EB8", manual = "#E41A1C")

# Set the position
ggplot(mtcars, aes(cyl, fill = am)) +
  geom_bar(position = 'dodge') +
  labs(x = "Number of Cylinders", y = "Count")
  scale_fill_manual("Transmission", values = palette)
```

# 2-8 Setting a dummy aesthetic


In the last chapter you saw that all the visible aesthetics can serve as attributes and aesthetics, but I very conveniently left out x and y. That's because although you can make univariate plots (such as histograms, which you'll get to in the next chapter), a y-axis will always be provided, even if you didn't ask for it.

You can make univariate plots in ggplot2, but you will need to add a fake y axis by mapping y to zero.

When using setting y-axis limits, you can specify the limits as separate arguments, or as a single numeric vector. That is, ylim(lo, hi) or ylim(c(lo, hi)).

```{r}
ggplot(mtcars, aes(mpg, 0)) +
  geom_jitter() +
  # Set the y-axis limits
  ylim(-2, 2)
```

## Geometries

# 3-1 Overplotting 1: large datasets

Scatter plots (using geom_point()) are intuitive, easily understood, and very common, but we must always consider overplotting, particularly in the following four situations:

Large datasets
Aligned values on a single axis
Low-precision data
Integer data
Typically, alpha blending (i.e. adding transparency) is recommended when using solid shapes. Alternatively, you can use opaque, hollow shapes.

Small points are suitable for large datasets with regions of high density (lots of overlapping).

Let's use the diamonds dataset to practice dealing with the large dataset case.


```{r}
# Plot price vs. carat, colored by clarity
plt_price_vs_carat_by_clarity <- ggplot(diamonds, aes(carat, price, color = clarity))

# Set transparency to 0.5
plt_price_vs_carat_by_clarity + geom_point(alpha = 0.5, shape = 16)
```

# 3-2 Overplotting 2: Aligned values

Let's take a look at another case where we should be aware of overplotting: Aligning values on a single axis.

This occurs when one axis is continuous and the other is categorical, which can be overcome with some form of jittering.

In the mtcars data set, fam and fcyl are categorical variants of cyl and am.


```{r}
# Plot base
plt_mpg_vs_fcyl_by_fam <- ggplot(mtcars, aes(cyl, mpg, color = am))

# Default points are shown for comparison
plt_mpg_vs_fcyl_by_fam + geom_point()

# Now jitter and dodge the point positions
plt_mpg_vs_fcyl_by_fam + geom_point(position=position_jitterdodge(jitter.width=0.3, dodge.width=0.3))


```

# 3-3 Overplotting 3: Low-precision data

You already saw how to deal with overplotting when using geom_point() in two cases:

Large datasets
Aligned values on a single axis
We used position = 'jitter' inside geom_point() or geom_jitter().

Let's take a look at another case:

Low-precision data
This results from low-resolution measurements like in the iris dataset, which is measured to 1mm precision (see viewer). It's similar to case 2, but in this case we can jitter on both the x and y axis.


```{r}
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  # Use a jitter position function with width 0.1
 # geom_point(alpha = 0.5)

 geom_point(position=position_jitter(width = 0.1), alpha = 0.5)
```

# 3-4 Overplotting 4: Integer data
Let's take a look at the last case of dealing with overplotting:

Integer data
This can be type integer (i.e. 1 ,2, 3…) or categorical (i.e. class factor) variables. factor is just a special class of type integer.

You'll typically have a small, defined number of intersections between two variables, which is similar to case 3, but you may miss it if you don't realize that integer and factor data are the same as low precision data.

The Vocab dataset provided contains the years of education and vocabulary test scores from respondents to US General Social Surveys from 1972-2004.


```{r}
library(carData)
# Examine the structure of Vocab
str(Vocab)

ggplot(Vocab, aes(education, vocabulary)) +
  # Set the shape to 1
  geom_jitter(alpha = 0.2, shape= 1)
```

# 3-5 Drawing histograms
Recall that histograms cut up a continuous variable into discrete bins and, by default, maps the internally calculated count variable (the number of observations in each bin) onto the y aesthetic. An internal variable called density can be accessed by using the .. notation, i.e. ..density... Plotting this variable will show the relative frequency, which is the height times the width of each bin.


```{r}
datacamp_light_blue <- "#51A8C9"

ggplot(mtcars, aes(mpg, ..density..)) +
  # Set the fill color to datacamp_light_blue
  geom_histogram(binwidth = 1, fill=datacamp_light_blue)
```

# 3-6 Positions in histograms
Here, we'll examine the various ways of applying positions to histograms. geom_histogram(), a special case of geom_bar(), has a position argument that can take on the following values:

stack (the default): Bars for different groups are stacked on top of each other.
dodge: Bars for different groups are placed side by side.
fill: Bars for different groups are shown as proportions.
identity: Plot the values as they appear in the dataset.


```{r}
ggplot(mtcars, aes(mpg, fill = am)) +
  # Change the position to identity, with transparency 0.4
  geom_histogram(binwidth = 1, position = "identity", alpha=0.4)
```

# 3-7 Position in bar and col plots
Let's see how the position argument changes geom_bar().

We have three position options:

stack: The default
dodge: Preferred
fill: To show proportions
While we will be using geom_bar() here, note that the function geom_col() is just geom_bar() where both the position and stat arguments are set to "identity". It is used when we want the heights of the bars to represent the exact values in the data.

In this exercise, you'll draw the total count of cars having a given number of cylinders (fcyl), according to manual or automatic transmission type (fam).


```{r}
ggplot(mtcars, aes(cyl, fill = am)) +
  # Change the position to "dodge"
  geom_bar(position = "dodge")
```

# 3-8 Overlapping bar plots
You can customize bar plots further by adjusting the dodging so that your bars partially overlap each other. Instead of using position = "dodge", you're going to use position_dodge(), like you did with position_jitter() in the the previous exercises. Here, you'll save this as an object, posn_d, so that you can easily reuse it.

Remember, the reason you want to use position_dodge() (and position_jitter()) is to specify how much dodging (or jittering) you want.

For this example, you'll use the mtcars dataset.


```{r}
ggplot(mtcars, aes(cyl, fill = am)) +
  # Set the transparency to 0.6
  geom_bar(position = position_dodge(width = 0.2), , alpha=0.6)
```

# 3-9 Bar plots: sequential color palette
In this bar plot, we'll fill each segment according to an ordinal variable. The best way to do that is with a sequential color palette.

Here's an example of using a sequential color palette with the mtcars dataset:

ggplot(mtcars, aes(fcyl, fill = fam)) +
  geom_bar() +
  scale_fill_brewer(palette = "Set1")
In the exercise, you'll use similar code on the the Vocab dataset. Both datasets are ordinal.


```{r}
# Plot education, filled by vocabulary
ggplot(Vocab, aes(education, fill = vocabulary)) +
  # Add a bar layer with position "fill"
  geom_bar(position = "fill") +
  # Add a brewer fill scale with default palette
  scale_fill_brewer(palette = "Set1")
```

# 3-10 Basic line plots
Here, we'll use the economics dataset to make some line plots. The dataset contains a time series for unemployment and population statistics from the Federal Reserve Bank of St. Louis in the United States. The data is contained in the ggplot2 package.

To begin with, you can look at how the median unemployment time and the unemployment rate (the number of unemployed people as a proportion of the population) change over time.


```{r}
# Change the y-axis to the proportion of the population that is unemployed
ggplot(economics, aes(date, unemploy/pop)) +
  geom_line()
```

# 3-11 Multiple time series
We already saw how the form of your data affects how you can plot it. Let's explore that further with multiple time series. Here, it's important that all lines are on the same scale, and if possible, on the same plot.

fish.species contains the global capture rates of seven salmon species from 1950–2010. Each variable (column) is a Salmon species and each observation (row) is one year. fish.tidy contains the same data, but in three columns: Species, Year, and Capture (i.e. one variable per column).


```{r}
#library(FSAdata)
#library(FSA)
# Plot the Rainbow Salmon time series
#ggplot(fish.species, aes(x = Year, y = Rainbow)) +
# geom_line()

# Plot the Pink Salmon time series
#ggplot(fish.species, aes(x = Year, y = Pink)) +
#  geom_line()

# Plot multiple time-series by grouping by species
#ggplot(fish.tidy, aes(Year, Capture)) +
#  geom_line(aes(group = Species))

# Plot multiple time-series by coloring by species
#ggplot(fish.tidy, aes(Year, Capture, color = Species)) +
#  geom_line()
```
## Themes

```{r}
plt_prop_unemployed_over_time = ggplot(economics, aes(x = date, y = unemploy/pop)) +
  ggtitle(c("The percentage of unemployed Americans \n increases sharply during recessions")) +
  geom_line()
```



# 4-1 Moving the legend
Let's wrap up this course by making a publication-ready plot communicating a clear message.

To change stylistic elements of a plot, call theme() and set plot properties to a new value. For example, the following changes the legend position.

p + theme(legend.position = new_value)

Here, the new value can be

"top", "bottom", "left", or "right'": place it at that side of the plot.
"none": don't draw it.
c(x, y): c(0, 0) means the bottom-left and c(1, 1) means the top-right.
Let's revisit the recession period line plot (assigned to plt_prop_unemployed_over_time).

```{r}
# Position the legend inside the plot at (0.6, 0.1)
plt_prop_unemployed_over_time +
  theme(legend.position=c(0.6, 0.1))
```

# 4-2 

```{r}
library(dplyr)
plt_prop_unemployed_over_time +
  theme(
    rect = element_rect(fill = "grey92"),
    legend.key = element_rect(color = NA),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    panel.grid.major.y = element_line(
      color = "white",
      size = 0.5,
      linetype = "dotted"
    ),
    # Set the axis text color to grey25
    axis.text=element_text(
      color = "grey25"),
    # Set the plot title font face to italic and font size to 16
   plot.title=element_text(
     size=16, face = "italic"
  ))
```

# 4-3 

```{r}
plt_mpg_vs_wt_by_cyl +
  theme(
    # Set the plot margin to (10, 30, 50, 70) millimeters
    plot.margin=margin(10, 30, 50, 70, "mm")
  )
```

# 4-4 

```{r}
# Add a void theme
plt_prop_unemployed_over_time +
 theme_void()
```

# 4-5 

```{r}
# Use the Wall Street Journal theme
plt_prop_unemployed_over_time +
  theme_wsj()
```

# 4-6 

```{r}
theme_recession <- theme(
  rect = element_rect(fill = "grey92"),
  legend.key = element_rect(color = NA),
  axis.ticks = element_blank(),
  panel.grid = element_blank(),
  panel.grid.major.y = element_line(color = "white", size = 0.5, linetype = "dotted"),
  axis.text = element_text(color = "grey25"),
  plot.title = element_text(face = "italic", size = 16),
  legend.position = c(0.6, 0.1)
)
theme_tufte_recession <- theme_tufte() + theme_recession

# Set theme_tufte_recession as the default theme
theme_set(theme_tufte_recession)

# Draw the plot (without explicitly adding a theme)
plt_prop_unemployed_over_time
```

# 4-7 

```{r}
plt_prop_unemployed_over_time +
  theme_tufte() +
  theme(
    legend.position = "none",
    axis.ticks = element_blank(),
    axis.title = element_text(color = "grey60"),
    axis.text = element_text(color = "grey60"),
    # Set the panel gridlines major y values
    panel.grid.major.y = element_line(
      # Set the color to grey60
     color="grey60",
      # Set the size to 0.25
      size=0.25,
      # Set the linetype to dotted
      linetype="dotted"
    )
  )
```

# 4-8 

```{r}
# Set the color scale
palette <- brewer.pal(5, "RdYlBu")[-(2:4)]

# Add a title and caption
ggplot(gm2007, aes(x = lifeExp, y = country, color = lifeExp)) +
  geom_point(size = 4) +
  geom_segment(aes(xend = 30, yend = country), size = 2) +
  geom_text(aes(label = round(lifeExp,1)), color = "white", size = 1.5) +
  scale_x_continuous("", expand = c(0,0), limits = c(30,90), position = "top") +
  scale_color_gradientn(colors = palette) +
  labs(title= "Highest and lowest life expectancies, 2007", caption = "Source: gapminder")
```

# 4-9 

```{r}
# Add a curve
plt_country_vs_lifeExp +  
  step_1_themes +
  geom_vline(xintercept = global_mean, color = "grey40", linetype = 3) +
  step_3_annotation +
  annotate(
    "curve",
    x = x_start, y = y_start,
    xend = x_end, yend = y_end,
    arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
    color = "grey40"
  )
```


