---
title: Machine Learning with Tree-Based Models in R
author: ''
date: '2021-07-26'
slug: []
categories: []
tags:
  - Tree-Based Models
  - Machine Learning
---

```{r}
library(tidymodels)
```

```{r}
# Load the package
library(tidymodels)
library(tidyverse)
diabetes <- read_csv("diabetes.csv")
```


## 1-3 Specify that tree

In order to build models and use them to solve real-world problems, you first need to lay the foundations of your model by creating a model specification. This is the very first step in every machine learning pipeline that you will ever build.

You are going to load the relevant packages and design the specification for your classification tree in just a few steps.

```{r}
# Pick a model class
tree_spec <- decision_tree() %>% 
  # Set the engine
  set_engine("rpart") %>% 
  # Set the mode
  set_mode("classification")

# Print the result
tree_spec
```

## 1-4 Train that model

A model specification is a good start, just like the canvas for a painter. But just as a painter needs color, the specification needs data. Only the final model is able to make predictions:

Model specification + data = model

In this exercise, you will train a decision tree that models the risk of diabetes using health variables as predictors. The response variable, outcome, indicates whether the patient has diabetes or not, which means this is a binary classification problem (there are just two classes). The dataset also contains health variables of patients like blood_pressure, age, and bmi.

```{r}
# Train the model
tree_model_bmi <- tree_spec %>% 
  fit(formula = factor(outcome) ~ bmi,
data = diabetes)

# Print the model
tree_model_bmi
```

## 1-6 Train/test split

In order to test your models, you need to build and test the model on two different parts of the data - otherwise, it's like cheating on an exam (as you already know the answers).

The data split is an integral part of the modeling process. You will dive into this by splitting the diabetes data and confirming the split proportions.

```{r}
# Create the split
diabetes_split <- initial_split(diabetes, prop = 0.8)

# Extract the training and test set
diabetes_train <- training(diabetes_split)
diabetes_test  <- testing(diabetes_split)

# Verify the proportions of both sets
round(nrow(diabetes_train) / nrow(diabetes), 2) == 0.80
round(nrow(diabetes_test) / nrow(diabetes), 2) == 0.20
```

## 1-7 Avoiding class imbalances

Some data contains very imbalanced outcomes - like a rare disease dataset. When splitting randomly, you might end up with a very unfortunate split. Imagine all the rare observations are in the test and none in the training set. That would ruin your whole training process!

Fortunately, the initial_split() function provides a remedy. You are going to observe and solve these so-called class imbalances in this exercise.

There's already a split object diabetes_split available with a 75% training and 25% test split.

```{r}
# Proportion of 'yes' outcomes in the training data
counts_train <- table(training(diabetes_split)$outcome)
prop_yes_train <- counts_train["yes"] / sum(counts_train)

# Proportion of 'yes' outcomes in the test data
counts_test <- table(testing(diabetes_split)$outcome)
prop_yes_test <- counts_test["yes"] / sum(counts_test)

paste("Proportion of positive outcomes in training set:", round(prop_yes_train, 2))
paste("Proportion of positive outcomes in test set:", round(prop_yes_test, 2))

```

```{r}
# Create a split with a constant outcome distribution
diabetes_split <- initial_split(diabetes,
prop = 0.75,
strata = outcome)

# Proportion of 'yes' outcomes in the training data
counts_train <- table(training(diabetes_split)$outcome)
prop_yes_train <- counts_train["yes"] / sum(counts_train)

# Proportion of 'yes' outcomes in the test data
counts_test <- table(testing(diabetes_split)$outcome)
prop_yes_test <- counts_test["yes"] / sum(counts_test)

paste("Proportion of positive outcomes in training set:", round(prop_yes_train, 2))
paste("Proportion of positive outcomes in test set:", round(prop_yes_test, 2))
```

## 1-8 From zero to hero

You mastered the skills of creating a model specification and splitting the data into training and test sets. You also know how to avoid class imbalances in the split. It's now time to combine what you learned in the preceding lesson and build your model using only the training set!

You are going to build a proper machine learning pipeline. This is comprised of creating a model specification, splitting your data into training and test sets, and last but not least, fitting the training data to a model. Enjoy!

```{r}
# Create the balanced data split
diabetes_split <- initial_split(diabetes,
prop = 0.75,
strata = outcome)

# Build the specification of the model
tree_spec <- decision_tree() %>% 
  set_engine("rpart") %>% 
  set_mode("classification")

# Train the model
model_trained <- tree_spec %>% 
  fit(formula = factor(outcome) ~ bmi+skin_thickness, 
      data = training(diabetes_split))

model_trained
```

## 1-10 Make predictions

Making predictions with data is one of the fundamental goals of machine learning. Now that you know how to split the data and fit a model, it's time to make predictions about unseen samples with your models.

You are going to make predictions about your test set using a model obtained by fitting the training data to a tree specification.

```{r}
model <- tree_spec %>% 
  fit(formula = factor(outcome) ~ ., 
      data = diabetes_train)

# Generate predictions
predictions <- predict(model,
                   diabetes_test, type = "class")

# Add the true outcomes
predictions_combined <- predictions %>% 
  mutate(true_class = diabetes_test$outcome)


# Print the first lines of the result
head(predictions_combined)
```

## 1-11 Crack the matrix

Visual representations are a great and intuitive way to assess results. One way to visualize and assess the performance of your model is by using a confusion matrix. In this exercise, you will create the confusion matrix of your predicted values to see in which cases it performs well and in which cases it doesn't.

```{r}
# The confusion matrix
diabetes_matrix <- conf_mat(data = predictions_combined,
estimate = .pred_class,
truth = true_class)

# Print the matrix
print(diabetes_matrix)
```

## 1-12 Are you predicting correctly?

Your model should be as good as possible, right? One way you can assess this is by counting how often it predicted the correct classes compared to the total number of predictions it made. As discussed in the video, we call this performance measure accuracy. You can either calculate this manually or by using a handy shortcut. Both obtain the same result.

```{r}

# The accuracy calculated by a function
acc_auto <- accuracy(predictions_combined, estimate = .pred_class, truth = factor(true_class))

acc_auto$estimate


```


##2-2 Train a regression tree

As you know already, decision trees are a useful tool for classification problems. Moreover, you can also use them to model regression problems. The structural difference is that there will be numeric values (instead of classes) on the leaf nodes.

In this exercise, you will use the chocolate dataset to fit a regression tree. This is very similar to what you already did in Chapter 1 with the diabetes dataset.

```{r}

chocolate_train <- read_csv("chocolate_train.csv")
# Build the specification
model_spec <- decision_tree() %>%
  set_mode("regression") %>%
  set_engine("rpart")

# Fit to the data
model_fit <- model_spec %>%
  fit(formula = final_grade ~ cocoa_percent+review_date,
      data = chocolate_train)

model_fit
```

##2-3 Predict new values

A predictive model is one that predicts the outcomes of new, unseen data. Besides the numeric predictors, there are other useful columns in the dataset. The goal of this exercise is to predict the final rating grades of a chocolate tasting based on all other predictor variables that are available.

```{r}
chocolate_test <- read_csv("chocolate_test.csv")

# Train the model
chocolate_model <- model_spec %>%
  fit(formula = final_grade ~ . , 
      data = chocolate_train)

# Predict new data
#predictions <- predict(chocolate_model,new_data = chocolate_test) %>%
  # Add the test set
 # bind_cols(chocolate_test)

#predictions
```

## 2-6 In-sample performance

It's very important to know whether your regression model is useful or not. A useful model can be one that captures the structure of your training set well. One way to assess this in-sample performance is to predict on training data and calculate the mean absolute error of all predicted data points.

In this exercise, you will evaluate your in-sample predictions using MAE (mean absolute error). MAE tells you approximately how far away the predictions are from the true values.