---
title: Machine Learning in the Tidyverse
author: ''
date: '2021-06-30'
slug: []
categories: []
tags:
  - data_wrangling
  - Machine Learning
---

## 1-2 Nesting your data

In this course, you will work with a collection of economic and social indicators for 77 countries over a period of 52 years. This data is stored in the gapminder dataframe.

In this exercise, you will transform your gapminder data into a nested dataframe by using the first tool needed to build the foundation of tidy machine learning skills: nest().

Note: This is a more granular version than the dataset available from the gapminder package. This version is available in the dslabs package.

```{r}
library(tidyverse)
gapminder <- read_csv("gapminder.csv")

# Explore gapminder
head(gapminder)

# Prepare the nested dataframe gap_nested

gap_nested <- gapminder %>% 
  group_by(country) %>% 
  nest()

# Explore gap_nested
head(gap_nested)
```

## 1-3 Unnesting your data

As you've seen in the previous exercise, a nested dataframe is simply a way to shape your data. Essentially taking the group_by() windows and packaging them in corresponding rows.

In the same way you can use the nest() function to break your data into nested chunks, you can use the unnest() function to expand the dataframes that are nested in these chunks.

```{r}
# Create the unnested dataframe called gap_unnnested
gap_unnested <- gap_nested %>% 
  unnest(col=data)
  
# Confirm that your data was not modified  
identical(gapminder, gap_unnested)
```

## 1-4 Explore a nested cell

In the first exercise, you successfully created a nested dataframe gap_nested. The data column contains tibbles for each country. In this exercise, you will explore one of these nested chunks.

```{r}
# Extract the data of Algeria
algeria_df <- gap_nested$data[[1]]

# Calculate the minimum of the population vector
min(algeria_df$population)

# Calculate the maximum of the population vector
max(algeria_df$population)

# Calculate the mean of the population vector
mean(algeria_df$population)
```

## 1-6 Mapping your data

In combination with mutate(), you can use map() to append the results of your calculation to a dataframe. Since the map() function always returns a vector of lists you must use unnest() to extract this information into a numeric vector.

Here you will explore this functionality by calculating the mean population of each country in the gapminder dataset. 

```{r}
# Calculate the mean population for each country
pop_nested <- gap_nested %>%
  mutate(mean_pop = map(data, ~mean(.x$population)))

# Take a look at pop_nested
head(pop_nested)

# Extract the mean_pop value by using unnest
pop_mean <- pop_nested %>% 
  unnest(mean_pop)

# Take a look at pop_mean
head(pop_mean)
```

## 1-7 

```{r}

```

## 1-8 

```{r}

```