---
title: Machine Learning in the Tidyverse
author: ''
date: '2021-06-30'
slug: []
categories: []
tags:
  - data_wrangling
  - Machine Learning
---

## 1-2 Nesting your data

In this course, you will work with a collection of economic and social indicators for 77 countries over a period of 52 years. This data is stored in the gapminder dataframe.

In this exercise, you will transform your gapminder data into a nested dataframe by using the first tool needed to build the foundation of tidy machine learning skills: nest().

Note: This is a more granular version than the dataset available from the gapminder package. This version is available in the dslabs package.

```{r}
library(tidyverse)
gapminder <- read_csv("gapminder.csv")

# Explore gapminder
head(gapminder)

# Prepare the nested dataframe gap_nested

gap_nested <- gapminder %>% 
  group_by(country) %>% 
  nest()

# Explore gap_nested
head(gap_nested)
```

## 1-3 Unnesting your data

As you've seen in the previous exercise, a nested dataframe is simply a way to shape your data. Essentially taking the group_by() windows and packaging them in corresponding rows.

In the same way you can use the nest() function to break your data into nested chunks, you can use the unnest() function to expand the dataframes that are nested in these chunks.

```{r}
# Create the unnested dataframe called gap_unnnested
gap_unnested <- gap_nested %>% 
  unnest(col=data)
  
# Confirm that your data was not modified  
identical(gapminder, gap_unnested)
```

## 1-4 Explore a nested cell

In the first exercise, you successfully created a nested dataframe gap_nested. The data column contains tibbles for each country. In this exercise, you will explore one of these nested chunks.

```{r}
# Extract the data of Algeria
algeria_df <- gap_nested$data[[1]]

# Calculate the minimum of the population vector
min(algeria_df$population)

# Calculate the maximum of the population vector
max(algeria_df$population)

# Calculate the mean of the population vector
mean(algeria_df$population)
```

## 1-6 Mapping your data

In combination with mutate(), you can use map() to append the results of your calculation to a dataframe. Since the map() function always returns a vector of lists you must use unnest() to extract this information into a numeric vector.

Here you will explore this functionality by calculating the mean population of each country in the gapminder dataset. 

```{r}
# Calculate the mean population for each country
pop_nested <- gap_nested %>%
  mutate(mean_pop = map(.x=data, .f=~mean(.x$population)))

# Take a look at pop_nested
head(pop_nested)

# Extract the mean_pop value by using unnest
pop_mean <- pop_nested %>% 
  unnest(mean_pop)

# Take a look at pop_mean
head(pop_mean)
```

## 1-7Expecting mapped output

When you know that the output of your mapped function is an expected type (here it is a numeric vector) you can leverage the map_*() family of functions to explicitly try to return that object type instead of a list.

Here you will again calculate the mean population of each country, but instead, you will use map_dbl() to explicitly append the numeric vector returned by mean() to your dataframe. 

```{r}
# Calculate mean population and store result as a double
pop_mean <- gap_nested %>%
  mutate(mean_pop = map_dbl(data, ~mean(.x$population)))

# Take a look at pop_mean
head(pop_mean)
```

## 1-8 Mapping many models

The gap_nested dataframe available in your workspace contains the gapminder dataset nested by country.

You will use this data to build a linear model for each country to predict life expectancy using the year feature.

Note: The term feature is synonymous with the terms variable or predictor. It refers to an attribute of your data that can be used to build a machine learning model.

```{r}
# Build a linear model for each country
gap_models <- gap_nested %>%
    mutate(model = map(data, ~lm(formula = life_expectancy~year, data = .x)))
    
# Extract the model for Algeria    
algeria_model <- gap_models$model[[1]]

# View the summary for the Algeria model
summary(algeria_model)
```


## 1-11 Extracting model statistics tidily

In this exercise, you will use the tidy() and glance() functions to extract information from algeria_model in a tidy manner.

For a linear model, tidy() extracts the model coefficients while glance() returns the model statistics such as the R^2.

```{r}
library(broom)

# Extract the coefficients of the algeria_model as a dataframe
tidy(algeria_model)

# Extract the statistics of the algeria_model as a dataframe
glance(algeria_model)
```

## 1-12 Augmenting your data

From the results of glance(), you learned that using the available features the linear model fits well with an adjusted  of 0.99. The augment() function can help you explore this fit by appending the predictions to the original data.

Here you will leverage this to compare the predicted values of life_expectancy with the original ones based on the year feature. 

```{r}
# Build the augmented dataframe
algeria_fitted <- augment(algeria_model)

# Compare the predicted values with the actual values of life expectancy
algeria_fitted %>% 
  ggplot(aes(x = year)) +
  geom_point(aes(y = life_expectancy)) + 
  geom_line(aes(y = .fitted), color = "red")
```

## 2-2 Tidy up the coefficients of your models

In this exercise you will leverage the list column workflow along with the tidy() function from broom to extract and explore the coefficients for the 77 models you built.

Remember the gap_models dataframe contains a model predicting life expectancy by year for 77 countries.

```{r}
# Extract the coefficient statistics of each model into nested dataframes
model_coef_nested <- gap_models %>% 
    mutate(coef = map(model, ~tidy(.x)))
    
# Simplify the coef dataframes for each model    
model_coef <- model_coef_nested %>%
    unnest(coef)

# Plot a histogram of the coefficient estimates for year         
model_coef %>% 
  filter(term == "year") %>% 
  ggplot(aes(x = estimate)) +
  geom_histogram()
```


## 2-5 Glance at the fit of your models

In this exercise you will use glance() to calculate how well the linear models fit the data for each country.

```{r}
# Extract the fit statistics of each model into dataframes
model_perf_nested <- gap_models %>% 
    mutate(fit = map(model, ~glance(.x)))

# Simplify the fit dataframes for each model    
model_perf <- model_perf_nested %>% 
    unnest(col=fit)
    
# Look at the first six rows of model_perf
head(model_perf)
```


## 2-6 Best and worst fitting models

In this exercise you will answer the following questions:

Overall, how well do your models fit your data?
Which are the best fitting models?
Which models do not fit the data well?

```{r}
# Plot a histogram of rsquared for the 77 models    
model_perf %>% 
  ggplot(aes(x = r.squared)) + 
  geom_histogram(bins = 30)  
  
# Extract the 4 best fitting models
best_fit <- model_perf %>% 
  top_n(n = 4, wt = r.squared)

# Extract the 4 models with the worst fit
worst_fit <- model_perf %>% 
  top_n(n = 4, wt = -r.squared)
```


## 2-8 Augment the fitted values of each model

In this exercise you will prepare your four best and worst fitting models for further exploration by augmenting your model data with augment().

```{r}
best_augmented <- best_fit %>% 
  # Build the augmented dataframe for each country model
  mutate(augmented = map(model, ~augment(.x))) %>% 
  # Expand the augmented dataframes
  unnest(augmented)

worst_augmented <- worst_fit %>% 
  # Build the augmented dataframe for each country model
  mutate(augmented = map(model, ~augment(.x))) %>% 
  # Expand the augmented dataframes
  unnest(augmented)
```


## 2-9 Explore your best and worst fitting models

Let's explore your four best and worst fitting models by comparing the fitted lines with the actual values.

```{r}
# Compare the predicted values with the actual values of life expectancy 
# for the top 4 best fitting models
best_augmented %>% 
  ggplot(aes(x = year)) +
  geom_point(aes(y = life_expectancy)) + 
  geom_line(aes(y = .fitted), color = "red") +
  facet_wrap(~country, scales = "free_y")
```

```{r}
# Compare the predicted values with the actual values of life expectancy 
# for the top 4 worst fitting models
worst_augmented %>% 
  ggplot(aes(x = year)) +
  geom_point(aes(y = life_expectancy)) + 
  geom_line(aes(y = .fitted), color = "red") +
  facet_wrap(~country, scales = "free_y")
```

## 2-11 Build better models
Earlier you built a collection of simple models to fit life expectancy using the year feature. Your previous analysis showed that some of these models didn't fit very well.

In this exercise you will build multiple regression models for each country using all available features. You may be interested in comparing the performance of the four worst fitting models so their adjusted  are provided below:

Country	Adjusted 
Botswana	-0.0060772
Lesotho	-0.0169851
Zambia	0.1668999
Zimbabwe	0.2083979

```{r}
# Build a linear model for each country using all features
gap_fullmodel <- gap_nested %>% 
  mutate(model = map(data, ~lm(life_expectancy~., data = .x)))

fullmodel_perf <- gap_fullmodel %>% 
  # Extract the fit statistics of each model into dataframes
  mutate(fit = map(model, ~glance(.x))) %>% 
  # Simplify the fit dataframes for each model
  unnest(fit)
  
# View the performance for the four countries with the worst fitting 
# four simple models you looked at before
fullmodel_perf %>% 
  filter(country %in% worst_fit$country) %>% 
  select(country, adj.r.squared)
```

## 3-2 The test-train split

In a disciplined machine learning workflow it is crucial to withhold a portion of your data (testing data) from any decision-making process. This allows you to independently assess the performance of your model when it is finalized. The remaining data, the training data, is used to build and select the best model.

In this exercise, you will use the rsample package to split your data to perform the initial train-test split of your gapminder data.

Note: Since this is a random split of the data it is good practice to set a seed before splitting it.

```{r}
library(rsample)
set.seed(42)

# Prepare the initial split object
gap_split <- initial_split(gapminder, prop = 0.75)

# Extract the training dataframe
training_data <- training(gap_split)

# Extract the testing dataframe
testing_data <- testing(gap_split)

# Calculate the dimensions of both training_data and testing_data
dim(training_data)
dim(testing_data)
```

## 3-3 Cross-validation dataframes

Now that you have withheld a portion of your data as testing data, you can use the remaining portion to find the best performing model.

In this exercise, you will split the training data into a series of 5 train-validate sets using the vfold_cv() function from the rsample package.

```{r}
set.seed(42)

# Prepare the dataframe containing the cross validation partitions
cv_split <- vfold_cv(training_data, v = 5)

cv_data <- cv_split %>% 
  mutate(
    # Extract the train dataframe for each split
    train = map(splits, ~training(.x)), 
    # Extract the validate dataframe for each split
    validate = map(splits, ~testing(.x))
  )

# Use head() to preview cv_data
head(cv_data)
```

## 3-5 Build cross-validated models

In this exercise, you will build a linear model predicting life_expectancy using all available features. You will do this for the train data of each cross-validation fold.

```{r}
# Build a model using the train data for each fold of the cross validation
cv_models_lm <- cv_data %>% 
  mutate(model = map(train, ~lm(formula = life_expectancy~., data = .x)))
```


## 3-6 Preparing for evaluation

In order to measure the validate performance of your models you need compare the predicted values of life_expectancy for the observations from validate set to the actual values recorded. Here you will prepare both of these vectors for each partition.


```{r}
cv_prep_lm <- cv_models_lm %>% 
  mutate(
    # Extract the recorded life expectancy for the records in the validate dataframes
    validate_actual = map(validate, ~.x$life_expectancy),
    # Predict life expectancy for each validate set using its corresponding model
    validate_predicted = map2(.x = model, .y = validate, ~predict(.x, .y))
  )
```

## 3-6 Evaluate model performance

Now that you have both the actual and predicted values of each fold you can compare them to measure performance.

For this regression model, you will measure the Mean Absolute Error (MAE) between these two vectors. This value tells you the average difference between the actual and predicted values.