---
title: Machine Learning in the Tidyverse
author: ''
date: '2021-06-30'
slug: []
categories: []
tags:
  - data_wrangling
  - Machine Learning
---

## 1-2 Nesting your data

In this course, you will work with a collection of economic and social indicators for 77 countries over a period of 52 years. This data is stored in the gapminder dataframe.

In this exercise, you will transform your gapminder data into a nested dataframe by using the first tool needed to build the foundation of tidy machine learning skills: nest().

Note: This is a more granular version than the dataset available from the gapminder package. This version is available in the dslabs package.

```{r}
library(tidyverse)
gapminder <- read_csv("gapminder.csv")

# Explore gapminder
head(gapminder)

# Prepare the nested dataframe gap_nested

gap_nested <- gapminder %>% 
  group_by(country) %>% 
  nest()

# Explore gap_nested
head(gap_nested)
```

## 1-3 Unnesting your data

As you've seen in the previous exercise, a nested dataframe is simply a way to shape your data. Essentially taking the group_by() windows and packaging them in corresponding rows.

In the same way you can use the nest() function to break your data into nested chunks, you can use the unnest() function to expand the dataframes that are nested in these chunks.

```{r}
# Create the unnested dataframe called gap_unnnested
gap_unnested <- gap_nested %>% 
  unnest(col=data)
  
# Confirm that your data was not modified  
identical(gapminder, gap_unnested)
```

## 1-4 Explore a nested cell

In the first exercise, you successfully created a nested dataframe gap_nested. The data column contains tibbles for each country. In this exercise, you will explore one of these nested chunks.

```{r}
# Extract the data of Algeria
algeria_df <- gap_nested$data[[1]]

# Calculate the minimum of the population vector
min(algeria_df$population)

# Calculate the maximum of the population vector
max(algeria_df$population)

# Calculate the mean of the population vector
mean(algeria_df$population)
```

## 1-6 Mapping your data

In combination with mutate(), you can use map() to append the results of your calculation to a dataframe. Since the map() function always returns a vector of lists you must use unnest() to extract this information into a numeric vector.

Here you will explore this functionality by calculating the mean population of each country in the gapminder dataset. 

```{r}
# Calculate the mean population for each country
pop_nested <- gap_nested %>%
  mutate(mean_pop = map(.x=data, .f=~mean(.x$population)))

# Take a look at pop_nested
head(pop_nested)

# Extract the mean_pop value by using unnest
pop_mean <- pop_nested %>% 
  unnest(mean_pop)

# Take a look at pop_mean
head(pop_mean)
```

## 1-7Expecting mapped output

When you know that the output of your mapped function is an expected type (here it is a numeric vector) you can leverage the map_*() family of functions to explicitly try to return that object type instead of a list.

Here you will again calculate the mean population of each country, but instead, you will use map_dbl() to explicitly append the numeric vector returned by mean() to your dataframe. 

```{r}
# Calculate mean population and store result as a double
pop_mean <- gap_nested %>%
  mutate(mean_pop = map_dbl(data, ~mean(.x$population)))

# Take a look at pop_mean
head(pop_mean)
```

## 1-8 Mapping many models

The gap_nested dataframe available in your workspace contains the gapminder dataset nested by country.

You will use this data to build a linear model for each country to predict life expectancy using the year feature.

Note: The term feature is synonymous with the terms variable or predictor. It refers to an attribute of your data that can be used to build a machine learning model.

```{r}
# Build a linear model for each country
gap_models <- gap_nested %>%
    mutate(model = map(data, ~lm(formula = life_expectancy~year, data = .x)))
    
# Extract the model for Algeria    
algeria_model <- gap_models$model[[1]]

# View the summary for the Algeria model
summary(algeria_model)
```


## 1-11 Extracting model statistics tidily

In this exercise, you will use the tidy() and glance() functions to extract information from algeria_model in a tidy manner.

For a linear model, tidy() extracts the model coefficients while glance() returns the model statistics such as the R^2.

```{r}
library(broom)

# Extract the coefficients of the algeria_model as a dataframe
tidy(algeria_model)

# Extract the statistics of the algeria_model as a dataframe
glance(algeria_model)
```

## 1-12 Augmenting your data

From the results of glance(), you learned that using the available features the linear model fits well with an adjusted  of 0.99. The augment() function can help you explore this fit by appending the predictions to the original data.

Here you will leverage this to compare the predicted values of life_expectancy with the original ones based on the year feature. 

```{r}
# Build the augmented dataframe
algeria_fitted <- augment(algeria_model)

# Compare the predicted values with the actual values of life expectancy
algeria_fitted %>% 
  ggplot(aes(x = year)) +
  geom_point(aes(y = life_expectancy)) + 
  geom_line(aes(y = .fitted), color = "red")
```

## 2-2 Tidy up the coefficients of your models

In this exercise you will leverage the list column workflow along with the tidy() function from broom to extract and explore the coefficients for the 77 models you built.

Remember the gap_models dataframe contains a model predicting life expectancy by year for 77 countries.

```{r}
# Extract the coefficient statistics of each model into nested dataframes
model_coef_nested <- gap_models %>% 
    mutate(coef = map(model, ~tidy(.x)))
    
# Simplify the coef dataframes for each model    
model_coef <- model_coef_nested %>%
    unnest(coef)

# Plot a histogram of the coefficient estimates for year         
model_coef %>% 
  filter(term == "year") %>% 
  ggplot(aes(x = estimate)) +
  geom_histogram()
```


## 2-5 

```{r}

```


## 2-6 

```{r}

```


## 2-8 

```{r}

```


## 2-9 

```{r}

```


## 2-11 

```{r}

```

