---
title: Writing Efficient R Code
author: ''
date: '2021-09-14'
slug: []
categories: []
tags:
  - R
---

## R version

One of the relatively easy optimizations available is to use an up-to-date version of R. In general, R is very conservative, so upgrading doesn't break existing code. However, a new version will often provide free speed boosts for key functions.

The version command returns a list that contains (among other things) the major and minor version of R currently being used.

```{r}
# Print the R version details using version
version

# Assign the variable major to the major component
major <- version$major

# Assign the variable minor to the minor component
minor <- version$minor
```
```{r}
library("rio")
movies_rds <- readRDS("movies.rds")
#export(movies_rds, "movies_csv.csv")
hh <- write_csv("movies_rds.rds")
```


## Comparing read times of CSV and RDS files

One of the most common tasks we perform is reading in data from CSV files. However, for large CSV files this can be slow. One neat trick is to read in the data and save as an R binary file (rds) using saveRDS(). To read in the rds file, we use readRDS().

Note: Since rds is R's native format for storing single objects, you have not introduced any third-party dependencies that may change in the future.

To benchmark the two approaches, you can use system.time(). This function returns the time taken to evaluate any R expression. For example, to time how long it takes to calculate the square root of the numbers from one to ten million, you would write the following:

system.time(sqrt(1:1e7))

```{r}
# How long does it take to read movies from CSV?
system.time(read.csv("movies.csv"))

# How long does it take to read movies from RDS?
system.time(readRDS("movies.rds"))
```
Using the <- operator inside a function call will create a new (or overwrite an existing) object.

## Elapsed time

Using system.time() is convenient, but it does have its drawbacks when comparing multiple function calls. The microbenchmark package solves this problem with the microbenchmark() function.


```{r warning=FALSE}
# Load the microbenchmark package
library(microbenchmark)

# Compare the two functions
compare <- microbenchmark(read.csv("movies.csv"), 
                          readRDS("movies.rds"), 
                          times = 10)

# Print compare
compare
```

