---
title: Writing Efficient R Code
author: ''
date: '2021-09-14'
slug: []
categories: []
tags:
  - R
---

## 1-2 R version

One of the relatively easy optimizations available is to use an up-to-date version of R. In general, R is very conservative, so upgrading doesn't break existing code. However, a new version will often provide free speed boosts for key functions.

The version command returns a list that contains (among other things) the major and minor version of R currently being used.

```{r}
# Print the R version details using version
version

# Assign the variable major to the major component
major <- version$major

# Assign the variable minor to the minor component
minor <- version$minor
```
```{r}
library("rio")
#movies_rds <- readRDS("movies.rds")
#export(movies_rds, "movies_csv.csv")
#hh <- readRDS("movies_rds.rds")
```


## 1-4 Comparing read times of CSV and RDS files

One of the most common tasks we perform is reading in data from CSV files. However, for large CSV files this can be slow. One neat trick is to read in the data and save as an R binary file (rds) using saveRDS(). To read in the rds file, we use readRDS().

Note: Since rds is R's native format for storing single objects, you have not introduced any third-party dependencies that may change in the future.

To benchmark the two approaches, you can use system.time(). This function returns the time taken to evaluate any R expression. For example, to time how long it takes to calculate the square root of the numbers from one to ten million, you would write the following:

system.time(sqrt(1:1e7))

```{r}
# How long does it take to read movies from CSV?
system.time(read.csv("movies.csv"))

# How long does it take to read movies from RDS?
system.time(readRDS("movies.rds"))
```
Using the <- operator inside a function call will create a new (or overwrite an existing) object.

##  1-6 Elapsed time

Using system.time() is convenient, but it does have its drawbacks when comparing multiple function calls. The microbenchmark package solves this problem with the microbenchmark() function.


```{r warning=FALSE}
# Load the microbenchmark package
library(microbenchmark)

# Compare the two functions
compare <- microbenchmark(read.csv("movies.csv"), 
                          readRDS("movies.rds"), 
                          times = 10)

# Print compare
compare
```

##  1-9 DataCamp hardware

For many problems your time is the expensive part. If having a faster computer makes you more productive, it can be cost effective to buy one. However, before you splash out on new toys for yourself, your boss/partner may want to see some numbers to justify the expense. Measuring the performance of your computer is called benchmarking, and you can do that with the benchmarkme package.

```{r}
# Load the benchmarkme package
library(benchmarkme)

# Assign the variable ram to the amount of RAM on this machine
ram <- get_ram()
ram

# Assign the variable cpu to the cpu specs
cpu <- get_cpu()
cpu
```

## 1-10 Benchmark DataCamp's machine
The benchmarkme package allows you to run a set of standardized benchmarks and compare your results to other users. One set of benchmarks tests is reading and writing speeds.

The function call

res = benchmark_io(runs = 1, size = 5)

records the length of time it takes to read and write a 5MB file.

```{r warning=FALSE}
# Load the package
library("benchmarkme")

# Run the benchmark
res <- benchmark_io(runs = 1, size = 5)

# Plot the results
plot(res)
```

## 2-3 Timings - growing a vector

Growing a vector is one of the deadly sins in R; you should always avoid it.

The growing() function defined below generates n random standard normal numbers, but grows the size of the vector each time an element is added!

Note: Standard normal numbers are numbers drawn from a normal distribution with mean 0 and standard deviation 1.
```{r}
#n <- 30000
# Slow code
growing <- function(n) {
    x <- NULL
    for(i in 1:n)
        x <- c(x, rnorm(1))
    x
}
```



```{r}
# Use <- with system.time() to store the result as res_grow
system.time(res_grow <- growing(30000))
```

## 2-4 Timings - pre-allocation
In the previous exercise, growing the vector took around 2 seconds. How long does it take when we pre-allocate the vector? The pre_allocate() function is defined below.

```{r}
#n <- 30000
# Fast code
pre_allocate <- function(n) {
    x <- numeric(n) # Pre-allocate
    for(i in 1:n) 
        x[i] <- rnorm(1)
    x
}
```



```{r}
# Use <- with system.time() to store the result as res_allocate
n <- 30000
system.time(res_allocate <- pre_allocate(n))
```

## 2-6 Vectorized code: multiplication

The following piece of code is written like traditional C or Fortran code. Instead of using the vectorized version of multiplication, it uses a for loop.

```{r}
x <- rnorm(10)
x2 <- numeric(length(x))
for(i in 1:10)
    x2[i] <- x[i] * x[i]

x
x2
```


Your job is to make this code more "R-like" by vectorizing it.

```{r}
# Store your answer as x2_imp
x2_imp <- x * x
```

## 2-7 Vectorized code: calculating a log-sum

A common operation in statistics is to calculate the sum of log probabilities. The following code calculates the log-sum (the sum of the logs).

# x is a vector of probabilities
total <- 0
for(i in seq_along(x)) 
    total <- total + log(x[i])
    
However this piece of code could be significantly improved using vectorized code.

```{r}
# Initial code
n <- 100
total <- 0
x <- runif(n)
for(i in 1:n) 
    total <- total + log(x[i])

# Rewrite in a single line. Store the result in log_sum
log_sum <- sum(log(x))
```

## 2-10 Data frames and matrices - column selection

All values in a matrix must have the same data type, which has efficiency implications when selecting rows and columns.

Suppose we have two objects, mat (a matrix) and df (a data frame).

---
# Which is faster, mat[, 1] or df[, 1]? 
microbenchmark(mat[, 1], df[, 1])
---

Q: What does selecting a row in a data frame return?

A: A data frame

## 2-12 Row timings

Similar to the previous example, the objects mat and df are a matrix and a data frame, respectively. Using microbenchmark(), how long does it take to select the first row from each of these objects?

To make the comparison fair, just use mat[1, ] and df[1, ].

---
# Which is faster, mat[1, ] or df[1, ]? 
microbenchmark(mat[1, ], df[1, ])
---

## 3-4 Profvis in action

Examine the code on the right that performs a standard data analysis. It loads and selects data, plots the data of interest, and adds in a regression line.

```{r}

library(ggplot2movies)
# Load the data set
data(movies, package = "ggplot2movies") 

# Load the profvis package
library(profvis)

# Profile the following code with the profvis function
profvis ({
  # Load and select data
  comedies <- movies[movies$Comedy == 1, ]

  # Plot data of interest
  plot(comedies$year, comedies$rating)

  # Loess regression line
  model <- loess(rating ~ year, data = comedies)
  j <- order(comedies$year)
  
  # Add fitted line to the plot
  lines(comedies$year[j], model$fitted[j], col = "red")
} )    ## Remember the closing brackets!
```

## 3-6 Change the data frame to a matrix

One of the parts of the code that profvis highlighted was the line where we generated the possible dice rolls and stored the results in a data frame:

df <- data.frame(d1 = sample(1:6, 3, replace = TRUE),
                d2 = sample(1:6, 3, replace = TRUE))
We can optimize this code by making two improvements:

Switching from a data frame to a matrix
Generating the 6 dice rolls in a single step
This gives

m <- matrix(sample(1:6, 6, replace = TRUE), ncol = 2)

```{r}
# Load the microbenchmark package
library(microbenchmark)

# The previous data frame solution is defined
# d() Simulates 6 dices rolls
d <- function() {
  data.frame(
    d1 = sample(1:6, 3, replace = TRUE),
    d2 = sample(1:6, 3, replace = TRUE)
  )
}

# Complete the matrix solution
m <- function() {
  matrix(sample(1:6, 3, replace = TRUE), ncol=2)
}

# Use microbenchmark to time m() and d()
microbenchmark(
 data.frame_solution = d(),
 matrix_solution     = m()
)
```

## 3-7 Calculating row sums

The second bottleneck identified was calculating the row sums.

total <- apply(d, 1, sum)
In the previous exercise you switched the underlying object to a matrix. This makes the above apply operation three times faster. But there's one further optimization you can use - switch apply() with rowSums().

```{r}
# Example data
s <- c(4,4,6,3,1,5)
rolls <- matrix(s, nrow = 3, ncol = 2)

# Define the previous solution 
app <- function(x) {
    apply(x, 1, sum)
}

# Define the new solution
r_sum <- function(x) {
    rowSums(x)
}

# Compare the methods
microbenchmark(
    app_sol = app(rolls),
    r_sum_sol = r_sum(rolls)
)
```

## 3-8 Use && instead of &

To determine if both dice are the same, the move_square() function uses if statements.

if (is_double[1] & is_double[2] & is_double[3]) {
    current <- 11 # Go To Jail - square 11 == Jail
}

The & operator will always evaluate both its arguments. That is, if you type x & y, R will always try to work out what x and y are. There are some cases where this is inefficient. For example, if x is FALSE, then x & y will always be FALSE, regardless of the value of y. Thus, you can save a little processing time by not calculating it. The && operator takes advantage of this trick, and doesn't bother to calculate y if it doesn't make a difference to the overall result.

In this code, if is_double[1] is FALSE we don't need to evaluate is_double[2] or is_double[3], so we can get a speedup by swapping & for &&.

One thing to note is that && only works on single logical values, i.e., logical vectors of length 1 (like you would pass into an if condition), but & also works on vectors of length greater than 1.

```{r}
# Example data
is_double <- c(FALSE, TRUE, TRUE)

# Define the previous solution
move <- function(is_double) {
    if (is_double[1] & is_double[2] & is_double[3]) {
        current <- 11 # Go To Jail
    }
}

# Define the improved solution
improved_move <- function(is_double) {
    if (is_double[1] && is_double[2] && is_double[3]) {
        current <- 11 # Go To Jail
    }
}

# microbenchmark both solutions
# Very occassionally the improved solution is actually a little slower
# This is just random chance
microbenchmark(move(is_double), improved_move(is_double), times = 1e5)
```

## 4-2 How many cores does this machine have?

The parallel package has a function detectCores() that determines the number of cores in a machine.

How many cores does this machine have?

```{r}
# Load the parallel package
library(parallel)

# Store the number of cores in the object no_of_cores
no_of_cores <- detectCores()

# Print no_of_cores
no_of_cores
```

## 4-8 Moving to parApply

To run code in parallel using the parallel package, the basic workflow has three steps.

Create a cluster using makeCluster().
Do some work.
Stop the cluster using stopCluster().
The simplest way to make a cluster is to pass a number to makeCluster(). This creates a cluster of the default type, running the code on that many cores.

The object dd is a data frame with 10 columns and 100 rows. The following code uses apply() to calculate the column medians:

apply(dd, 2, median)
To run this in parallel, you swap apply() for parApply(). The arguments to this function are the same, except that it takes a cluster argument before the usual apply() arguments.

```{r}

```

## 4-10 Using parSapply()

We previously played the following game:

Initialize: total = 0.
Roll a single die and add it to total.
If total is even, reset total to zero.
If total is greater than 10. The game finishes.
The game could be simulated using the play() function:

play <- function() {
  total <- no_of_rolls <- 0
  while(total < 10) {
    total <- total + sample(1:6, 1)

    # If even. Reset to 0
    if(total %% 2 == 0) total <- 0 
    no_of_rolls <- no_of_rolls + 1
  }
  no_of_rolls
}
To simulate the game 100 times, we could use a for loop or sapply():

```{r, eval=FALSE} res <- sapply(1:100, function(i) play())

This is perfect for running in parallel!

To make functions available on a cluster, you use the [`clusterExport()`](https://www.rdocumentation.org/packages/parallel/topics/clusterApply) function. This takes a cluster and a string naming the function.
{r} clusterExport(cl, "some_function") ```

```{r}

```

## 4-11 Timings parSapply()

Running the dice game is embarrassingly parallel. These types of simulations usually (but not always) produce a good speed-up. As before, we can use microbenchmark() or system.time(). For simplicity, we'll use system.time() in this exercise.

```{r}

```

