---
title: Writing Efficient R Code
author: ''
date: '2021-09-14'
slug: []
categories: []
tags:
  - R
---

## 1-2 R version

One of the relatively easy optimizations available is to use an up-to-date version of R. In general, R is very conservative, so upgrading doesn't break existing code. However, a new version will often provide free speed boosts for key functions.

The version command returns a list that contains (among other things) the major and minor version of R currently being used.

```{r}
# Print the R version details using version
version

# Assign the variable major to the major component
major <- version$major

# Assign the variable minor to the minor component
minor <- version$minor
```
```{r}
library("rio")
movies_rds <- readRDS("movies.rds")
#export(movies_rds, "movies_csv.csv")
hh <- write_csv("movies_rds.rds")
```


## 1-4 Comparing read times of CSV and RDS files

One of the most common tasks we perform is reading in data from CSV files. However, for large CSV files this can be slow. One neat trick is to read in the data and save as an R binary file (rds) using saveRDS(). To read in the rds file, we use readRDS().

Note: Since rds is R's native format for storing single objects, you have not introduced any third-party dependencies that may change in the future.

To benchmark the two approaches, you can use system.time(). This function returns the time taken to evaluate any R expression. For example, to time how long it takes to calculate the square root of the numbers from one to ten million, you would write the following:

system.time(sqrt(1:1e7))

```{r}
# How long does it take to read movies from CSV?
system.time(read.csv("movies.csv"))

# How long does it take to read movies from RDS?
system.time(readRDS("movies.rds"))
```
Using the <- operator inside a function call will create a new (or overwrite an existing) object.

##  1-6 Elapsed time

Using system.time() is convenient, but it does have its drawbacks when comparing multiple function calls. The microbenchmark package solves this problem with the microbenchmark() function.


```{r warning=FALSE}
# Load the microbenchmark package
library(microbenchmark)

# Compare the two functions
compare <- microbenchmark(read.csv("movies.csv"), 
                          readRDS("movies.rds"), 
                          times = 10)

# Print compare
compare
```

##  1-9 DataCamp hardware

For many problems your time is the expensive part. If having a faster computer makes you more productive, it can be cost effective to buy one. However, before you splash out on new toys for yourself, your boss/partner may want to see some numbers to justify the expense. Measuring the performance of your computer is called benchmarking, and you can do that with the benchmarkme package.

```{r}
# Load the benchmarkme package
library(benchmarkme)

# Assign the variable ram to the amount of RAM on this machine
ram <- get_ram()
ram

# Assign the variable cpu to the cpu specs
cpu <- get_cpu()
cpu
```

## 1-10 Benchmark DataCamp's machine
The benchmarkme package allows you to run a set of standardized benchmarks and compare your results to other users. One set of benchmarks tests is reading and writing speeds.

The function call

res = benchmark_io(runs = 1, size = 5)

records the length of time it takes to read and write a 5MB file.

```{r warning=FALSE}
# Load the package
library("benchmarkme")

# Run the benchmark
res <- benchmark_io(runs = 1, size = 5)

# Plot the results
plot(res)
```

