---
title: Dealing With Missing Data in R
author: ''
date: '2021-09-06'
slug: []
categories: []
tags:
  - data_wrangling
---

```{r}
library(naniar)

dat_hw <- read.csv("dat_hw.csv")
```

## 1-2 Using and finding missing values

When working with missing data, there are a couple of commands that you should be familiar with - firstly, you should be able to identify if there are any missing values, and where these are.

Using the any_na() and are_na() tools, identify which values are missing.

```{r}
# Create x, a vector, with values NA, NaN, Inf, ".", and "missing"
x <- c(NA, NaN, Inf, ".", "missing")

# Use any_na() and are_na() on to explore the missings
any_na(x)
are_na(x)
```

## 1-3 How many missing values are there?

One of the first things that you will want to check with a new dataset is if there are any missing missing values, and how many there are.

You could use are_na() to and count up the missing values, but the most efficient way to count missings is to use the n_miss() function. This will tell you the total number of missing values in the data.

You can then find the percent of missing values in the data with the pct_miss function. This will tell you the percentage of missing values in the data.

You can also find the complement to these - how many complete values there are - using n_complete and pct_complete.

```{r}
# Use n_miss() to count the total number of missing values in dat_hw
n_miss(dat_hw)

# Use n_miss() on dat_hw$weight to count the total number of missing values
n_miss(dat_hw$weight)

# Use n_complete() on dat_hw to count the total number of complete values
n_complete(dat_hw)

# Use n_complete() on dat_hw$weight to count the total number of complete values
n_complete(dat_hw$weight)

# Use prop_miss() and prop_complete() on dat_hw to count the total number of missing values in each of the variables
prop_miss(dat_hw)
prop_complete(dat_hw)
```

## 1-6 Summarizing missingness

Now that you understand the behavior of missing values in R, and how to count them, let's scale up our summaries for cases (rows) and variables, using miss_var_summary() and miss_case_summary(), and also explore how they can be applied for groups in a dataframe, using the group_by function from dplyr.

```{r}

```

## 1-7 

```{r}

```

## 1-8 

```{r}

```

## 1-10 

```{r}

```

## 1-11 

```{r}

```

## 1-12 

```{r}

```

